---
title: "NACI_PaperII"
author: "Tommy J. Flynn"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    bibliography: paperII_citations.bib
---

```{r setup}
library(tidyverse)
library(lubridate)
# library(naniar) # for missing data
library(vtable) # makes good summary/descriptive statistics table
library(kableExtra)
library(RefManageR)
library(readxl)
library(haven)
# set knit options
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      collapse = TRUE)

# create path to data directory
rdata_dir <- "/Users/tommy-two/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data/Rdata"
```


# Code

__Population from all shifts during the observation year (2009-2010):__

```{r population_730shifts}
# load the population data from the full year, all 730 shifts"
population_730 <- read_rds(paste0(rdata_dir, "/population.Rdata"))
# clean race variable for analysis
population_730 <- population_730 %>%
  mutate(race = factor(race,
              levels = c("Black",
                         "White",
                         "Other"),
              exclude = c("Not Recorded", "1", "2", "3")))

population_730 %>% 
  count(race) %>% 
  mutate(`%` = round(n/sum(n)*100, 2)) %>% 
  kable(caption = "Distribution of self-identified race for entire ED patient population from the study year")

# check results
# population_730 %>% 
  # glimpse()
# how many patients in population from 2009-2010?
n_population_730 <- population_730 %>%
  nrow()
freq_flyer_nums <- population_730 %>% 
  count(med_rec_no) %>%
  filter(n > 1) %>%
  arrange(desc(n))
  # mutate(med_rec_size = str_length(med_rec_no)) %>% 
  # summarise(max_lenght = max(med_rec_size), min_length = min(med_rec_size), median_length = median(med_rec_size),
  #           mean_length = mean(med_rec_size))
  # med_rec_no is 5 to 7 characters in length, with a median of 7 and mean of 6.95. How many med_rec_no have length 5 = 156, over 2,000 have 6 characters. 
n_freq_flyers <- freq_flyer_nums %>% 
  nrow()
n_freq_flyer_visits <- freq_flyer_nums %>% 
  summarise(freq_flyer_visits = sum(n)) %>% 
  pull(freq_flyer_visits)
  # filter(n >= 3) %>% 
  # summarise(mean_visits = mean(n), median_visits = median(n), sd_visits = sd(n), iqr_visits = IQR(n))
n_unique_pats730 <- population_730 %>% 
  distinct(med_rec_no) %>% 
  nrow()
```

__Population from 81 shifts with observational data:__

```{r population_81shifts}
# load population data from the 81 observed shifts"
population_81 <- read_rds(paste0(rdata_dir, "/pt_info_final.Rdata"))

population_81 <- population_81 %>% 
  mutate(race = factor(race,
              levels = c("Black",
                         "White",
                         "Other"),
              exclude = c("Not Recorded", "1", "2", "3")))
# how many patients in population of 81 shifts?
n_population_81 <- population_81 %>%
  nrow()
# n_population_81
population_81 %>% 
  count(race) %>%
  mutate(`%` = round(n/sum(n)*100, 2)) %>% 
  kable(caption = "Distribution of self-identified race for ED patient population during 81 shifts")
```


```{r shift_nums}
# pull a column of all unique shift numbers
shift_nums <- population_81 %>% 
  distinct(shift_num)

# create a vector of all shift numbers from that column
shift_nums_vec <- pull(shift_nums, shift_num)
rm(shift_nums)
```


```{r shift_dates}
# create a variable with date intervals for patients' ED stays from arrival date and departure date
pop_730_interval <- population_730 %>% 
  mutate(ed_interval = interval(arrival_date, 
                                departure_date),
         los_seconds = int_length(ed_interval))


pop_730_interval %>% count(los) %>% 
  arrange(desc(n)) %>% 
  mutate(los_seconds = seconds(los),
         los_days = days(los_seconds))



pop_730_interval %>% 
  mutate(shift_ampm = if_else(as_date("2009-07-09") %within% ed_interval, "pm", NA_character_)) %>% 
  count(shift_ampm)
population_81 %>% filter(as_date(arrival_dttm) == "2009-07-09" | as_date(departure_dttm) == "2009-07-09")

population_81 %>% mutate(los = NULL,
                         job = "PATIENT")
  


# shift_dates <- 

population_81 %>% 
  distinct(date, shift_num)
```


```{r shift_subsample}
# set seed for reproducible random sampling
set.seed(711)
# randomly sample 20 shifts
shift_sample_vec <- sample(shift_nums_vec, 20)

population_20 <- population_730 %>% 
  filter(shift_num %in% shift_sample_vec) 
# number of patients in population_20
n_population_20 <- population_20 %>% 
  nrow()
pop_20_arrdate <- population_20 %>% 
  mutate(arrival_date = date(arrival_dttm)) %>% 
  pull(arrival_date)
population_20 %>% count(shift_num) %>% 
  arrange(n)
names(population_81)
```



```{r sample_81}
# load data for patients included in the study:
sample_81 <- read_rds(paste0(rdata_dir, "/pts_all.Rdata"))
# wrangle sample_81 to exclude unknown/not recorded/or undefined observations:
sample_81 <- sample_81 %>% 
  mutate(participant_cat = replace_na(job, "PATIENT"), 
         race = factor(race,
              levels = c("Black",
                         "White",
                         "Other"),
              exclude = c("Not Recorded", "1", "2", "3")),
         arrival_mode = na_if(arrival_mode, "Unknown"),
         disposition = na_if(disposition, "Not Recorded"),
         .keep = "unused") %>%
  relocate(participant_cat, .after = pt_id)
# check results:
# sample_81 %>% 
#   glimpse()
sample_81 %>% 
  # filter(participant == "Yes") %>%
  # nrow()
  count(race) %>% 
  mutate(`%` = round(n/sum(n)*100, 2)) %>% 
  kable(caption = "Distribution of self-identified race for ED patient sample from 81 shifts")
# how many patient participants?
n_sample_81 <- sample_81 %>% 
  nrow()
# n_sample_81
sample_20 <-   sample_81 %>% 
  filter(shift_num %in% shift_sample_vec)

# sample_81_20 %>% 
  # nrow()
  # select(-nested_loc) %>% 
  # filter(!is.na(participant) & !is.na(approached) & !is.na(participant_final)) %>% # NA x 443
  # select(race,
  #        age,
  #        sex,
  #        acuity, 
  #        arrival_cat,
  #        los_hours,) %>% 
  # vtable::st()

# how many patients are in the 20 sampled shifts?
n_sample_20 <- sample_20 %>% 
  nrow()
# what are their races?
sample_20 %>% 
  count(race) %>% 
  mutate(`%` = round(n/sum(n)*100, 2))%>% 
  kable(caption = "Distribution of self-identified race for ED patient sample from 20 shifts")
# what was their sex?
sample_20 %>% 
  count(sex) %>% 
  mutate(`%` = round(n/sum(n)*100, 2)) %>% 
  kable(caption = "Sex of ED patients from 20 shifts")
# what was their age?
sample_20 %>% 
  filter(!is.na(age)) %>% 
  summarise(age_mean = mean(age), 
            sd_age = sd(age),
            age_min = min(age), 
            age_max = max(age))%>% 
  kable(caption = "Distribution of patient age for ED patient sample from 20 shifts")
```








```{r staff_exploration}
staff_all <- read_rds(paste0(rdata_dir, "/staff_all.Rdata"))
# pull data for staff observed during the 20 sampled shifts
staff_sample20 <- staff_all %>% 
  filter(shift_num %in% shift_nums_vec)
# group by sid to count individual staff
n_staff_sample <- staff_sample20 %>% 
  ungroup() %>% 
  distinct(sid) %>% 
  nrow()

# glimpse(staff_sample)


```






# Patient Population and Sample

Demographics for the whole ED patient population in the year in which data were collected were reported previously [@lowery-north2013]. In summary, `r n_unique_pats730` unique patients visited the the ED `r n_population_730` times between July 1, 2009 and June 30, 2010. `r n_freq_flyers` (`r round(n_freq_flyers/n_unique_pats730 * 100, 1)`%) patients visited the ED more than once and were collectively responsible for `r n_freq_flyer_visits` (`r round(n_freq_flyer_visits/n_population_730 * 100, 1)`%). encounters.
`r n_population_81` (`r round(n_population_81/n_population_730*100, 1)`%) visits occurred during 81 observed shifts.

Of patients that visited the ED during these 81 observed shifts, 34% (3112) were not approached by the research team, 16% (941) were excluded for reasons related to the patient (e.g., refused or unable to consent), and 6% (389) had technical issues which precluded inclusion. Data were, therefore, available for `r n_sample_81` (`r round(n_sample_81/n_population_81*100, 1)`%) patients from 81 shifts, of which `r n_sample_20` (`r round(n_sample_20/n_sample_81*100, 1)`%).

# Staff Sample

`r n_staff_sample` (`r round(n_staff_sample/104*100, 1)`%) of 104 eligible staff consented to participate [@lowery-north2013], and all `r n_staff_sample` (100%) of those staff members were present during at least one of the 20 shifts included in this secondary analysis.

