---
title: "Networks by Shift"
author: "Tommy J. Flynn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
library(igraph)
library(tidyverse)
library(tsna)
library(ndtv)
library(animation)
library(networkDynamic)
```


```{r, warning=FALSE}
layouts <- grep("^layout_.", ls(package:igraph), value=TRUE)
# Remove layouts that do not apply to our graph (not bipartite)
layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree", layouts)]

layouts
```



# Shift 1


```{r}
shift_1 <- read_rds("Data/networks/shift_1_net.Rdata")
net_1 <- graph_from_data_frame(d= shift_1$edges, vertices = shift_1$vertices, directed=F)
# calculate degree and resize vertices 
# deg <- degree(net_1, mode = "total")
# V(net_1)$size <- deg

# plot(net_1)



# set edge width equal to colocation duration in hours
# E(net_1)$width <- E(net_1)$coloc_duration/60/60

# create examples of all relevant layouts in igraph
par(mfrow=c(3,3), mar=c(1,1,1,1))
for (layout in layouts) {
  print(layout)
  l <- do.call(layout, list(net_1))
  plot(net_1, layout=l, main=layout) 
}

library(ggraph)
library(ggplot2)
ggraph(net_1, layout = "fr") +
  geom_edge_link() + 
  geom_node_text(aes(label = sid_type)) +
  ggtitle("Shift 1") + 
  theme_void()
```


# Shift 2

```{r}
shift_8 <- read_rds('Data/networks/shift_8_net.Rdata')
net_8 <- graph_from_data_frame(d= shift_8$edges, vertices = shift_8$vertices, directed=F)
plot(net_8)

```

# Shift 10

```{r}
shift_10 <- read_rds('Data/networks/shift_10_net.Rdata')
net_10 <- graph_from_data_frame(d= shift_10$edges, vertices = shift_10$vertices, directed=F)
plot(net_10)


```

# Shift 17

```{r}
shift_17 <- read_rds('Data/networks/shift_17_net.Rdata')
net_17 <- graph_from_data_frame(d= shift_17$edges, vertices = shift_17$vertices, directed=F)
plot(net_17)

```

# Shift 19

```{r}
shift_19 <- read_rds('Data/networks/shift_19_net.Rdata')
net_19 <- graph_from_data_frame(d= shift_19$edges, vertices = shift_19$vertices, directed=F)
plot(net_19)
```

# Shift 23

```{r}
shift_23 <- read_rds('Data/networks/shift_23_net.Rdata')
net_23 <- graph_from_data_frame(d= shift_23$edges, vertices = shift_23$vertices, directed=F)
plot(net_23)
```

# Shift 25

```{r}
shift_25 <- read_rds('Data/networks/shift_25_net.Rdata')
net_25 <- graph_from_data_frame(d= shift_25$edges, vertices = shift_25$vertices, directed=F)
plot(net_25)

```

# Shift 38

```{r}
shift_38 <- read_rds('Data/networks/shift_38_net.Rdata')
net_38 <- graph_from_data_frame(d= shift_38$edges, vertices = shift_38$vertices, directed=F)
plot(net_38)


```

# Shift 43

```{r}
shift_43 <- read_rds('Data/networks/shift_43_net.Rdata')
net_43 <- graph_from_data_frame(d= shift_43$edges, vertices = shift_43$vertices, directed=F)
plot(net_43)
```

# Shift 47

```{r}
shift_47 <- read_rds('Data/networks/shift_47_net.Rdata')
net_47 <- graph_from_data_frame(d= shift_47$edges, vertices = shift_47$vertices, directed=F)
plot(net_47)
```

# Shift 53

```{r}
shift_53 <- read_rds('Data/networks/shift_53_net.Rdata')
net_53 <- graph_from_data_frame(d= shift_53$edges, vertices = shift_53$vertices, directed=F)
plot(net_53)


```

# Shift 56

```{r}
shift_56 <- read_rds('Data/networks/shift_56_net.Rdata')
net_56 <- graph_from_data_frame(d= shift_56$edges, vertices = shift_56$vertices, directed=F)
plot(net_56)


```

# Shift 63

```{r}
shift_63 <- read_rds('Data/networks/shift_63_net.Rdata')
net_63 <- graph_from_data_frame(d= shift_63$edges, vertices = shift_63$vertices, directed=F)
plot(net_63)


```

# Shift 67

```{r}
shift_67 <- read_rds('Data/networks/shift_67_net.Rdata')
net_67 <- graph_from_data_frame(d= shift_67$edges, vertices = shift_67$vertices, directed=F)
plot(net_67)


```

# Shift 78

```{r}
shift_78 <- read_rds('Data/networks/shift_78_net.Rdata')
net_78 <- graph_from_data_frame(d= shift_78$edges, vertices = shift_78$vertices, directed=F)
plot(net_78)


```

# Shift 81

```{r}
shift_81 <- read_rds('Data/networks/shift_81_net.Rdata')
net_81 <- graph_from_data_frame(d= shift_81$edges, vertices = shift_81$vertices, directed=F)
plot(net_81)


```

# Shift 86

```{r}
shift_86 <- read_rds('Data/networks/shift_86_net.Rdata')
net_86 <- graph_from_data_frame(d= shift_86$edges, vertices = shift_86$vertices, directed=F)
plot(net_86)


```

# Shift 91

```{r}
shift_91 <- read_rds('Data/networks/shift_91_net.Rdata')
net_91 <- graph_from_data_frame(d= shift_91$edges, vertices = shift_91$vertices, directed=F)
plot(net_91)

```

# Shift 93

```{r}
shift_93 <- read_rds('Data/networks/shift_93_net.Rdata')
net_93 <- graph_from_data_frame(d= shift_93$edges, vertices = shift_93$vertices, directed=F)
plot(net_93)
```

# Shift 96

```{r}
shift_96 <- read_rds('Data/networks/shift_96_net.Rdata')
net_96 <- graph_from_data_frame(d= shift_96$edges, vertices = shift_96$vertices, directed=F)
plot(net_96)

```

# Shift 98

```{r}
shift_98 <- read_rds('Data/networks/shift_98_net.Rdata')
net_98 <- graph_from_data_frame(d= shift_98$edges, vertices = shift_98$vertices, directed=F)
plot(net_98)
```

# Shift 102

```{r}
shift_102 <- read_rds('Data/networks/shift_102_net.Rdata')
net_102 <- graph_from_data_frame(d= shift_102$edges, vertices = shift_102$vertices, directed=F)
plot(net_102)
```

# Shift 104

```{r}
shift_104 <- read_rds('Data/networks/shift_104_net.Rdata')
net_104 <- graph_from_data_frame(d= shift_104$edges, vertices = shift_104$vertices, directed=F)
plot(net_104)

```

# Shift 112

```{r}
shift_112 <- read_rds('Data/networks/shift_112_net.Rdata')
net_112 <- graph_from_data_frame(d= shift_112$edges, vertices = shift_112$vertices, directed=F)
plot(net_112)

```

# Shift 114

```{r}
shift_114 <- read_rds('Data/networks/shift_114_net.Rdata')
net_114 <- graph_from_data_frame(d= shift_114$edges, vertices = shift_114$vertices, directed=F)
plot(net_114)

```

# Shift 132

```{r}
shift_132 <- read_rds('Data/networks/shift_132_net.Rdata')
net_132 <- graph_from_data_frame(d= shift_132$edges, vertices = shift_132$vertices, directed=F)
plot(net_132)

```

# Shift 138

```{r}
shift_138 <- read_rds('Data/networks/shift_138_net.Rdata')
net_138 <- graph_from_data_frame(d= shift_138$edges, vertices = shift_138$vertices, directed=F)
plot(net_138)

```

# Shift 145

```{r}
shift_145 <- read_rds('Data/networks/shift_145_net.Rdata')
net_145 <- graph_from_data_frame(d= shift_145$edges, vertices = shift_145$vertices, directed=F)
plot(net_145)

```

# Shift 148

```{r}
shift_148 <- read_rds('Data/networks/shift_148_net.Rdata')
net_148 <- graph_from_data_frame(d= shift_148$edges, vertices = shift_148$vertices, directed=F)
plot(net_148)
```

# Shift 155

```{r}
shift_155 <- read_rds('Data/networks/shift_155_net.Rdata')
net_155 <- graph_from_data_frame(d= shift_155$edges, vertices = shift_155$vertices, directed=F)
plot(net_155)


```

# Shift 161

```{r}
shift_161 <- read_rds('Data/networks/shift_161_net.Rdata')
net_161 <- graph_from_data_frame(d= shift_161$edges, vertices = shift_161$vertices, directed=F)
plot(net_161)

```

# Shift 163

```{r}
shift_163 <- read_rds('Data/networks/shift_163_net.Rdata')
net_163 <- graph_from_data_frame(d= shift_163$edges, vertices = shift_163$vertices, directed=F)
plot(net_163)

```

# Shift 166

```{r}
shift_166 <- read_rds('Data/networks/shift_166_net.Rdata')
net_166 <- graph_from_data_frame(d= shift_166$edges, vertices = shift_166$vertices, directed=F)
plot(net_166)

```

# Shift 168

```{r}
shift_168 <- read_rds('Data/networks/shift_168_net.Rdata')
net_168 <- graph_from_data_frame(d= shift_168$edges, vertices = shift_168$vertices, directed=F)
plot(net_168)

```

# Shift 172

```{r}
shift_172 <- read_rds('Data/networks/shift_172_net.Rdata')
net_172 <- graph_from_data_frame(d= shift_172$edges, vertices = shift_172$vertices, directed=F)
plot(net_172)

```

# Shift 179

```{r}
shift_179 <- read_rds('Data/networks/shift_179_net.Rdata')
net_179 <- graph_from_data_frame(d= shift_179$edges, vertices = shift_179$vertices, directed=F)
plot(net_179)

```

# Shift 181

```{r}
shift_181 <- read_rds('Data/networks/shift_181_net.Rdata')
net_181 <- graph_from_data_frame(d= shift_181$edges, vertices = shift_181$vertices, directed=F)
plot(net_181)

```

# Shift 189

```{r}
shift_189 <- read_rds('Data/networks/shift_189_net.Rdata')
net_189 <- graph_from_data_frame(d= shift_189$edges, vertices = shift_189$vertices, directed=F)
plot(net_189)

```

# Shift 191

```{r}
shift_191 <- read_rds('Data/networks/shift_191_net.Rdata')
net_191 <- graph_from_data_frame(d= shift_191$edges, vertices = shift_191$vertices, directed=F)
plot(net_191)
```

# Shift 195

```{r}
shift_195 <- read_rds('Data/networks/shift_195_net.Rdata')
net_195 <- graph_from_data_frame(d= shift_195$edges, vertices = shift_195$vertices, directed=F)
plot(net_195)


```

# Shift 197

```{r}
shift_197 <- read_rds('Data/networks/shift_197_net.Rdata')
net_197 <- graph_from_data_frame(d= shift_197$edges, vertices = shift_197$vertices, directed=F)
plot(net_197)

```

# Shift 201

```{r}
shift_201 <- read_rds('Data/networks/shift_201_net.Rdata')
net_201 <- graph_from_data_frame(d= shift_201$edges, vertices = shift_201$vertices, directed=F)
plot(net_201)

```

# Shift 206

```{r}
shift_206 <- read_rds('Data/networks/shift_206_net.Rdata')
net_206 <- graph_from_data_frame(d= shift_206$edges, vertices = shift_206$vertices, directed=F)
plot(net_206)

```

# Shift 209

```{r}
shift_209 <- read_rds('Data/networks/shift_209_net.Rdata')
net_209 <- graph_from_data_frame(d= shift_209$edges, vertices = shift_209$vertices, directed=F)
plot(net_209)

```

# Shift 214

```{r}
shift_214 <- read_rds('Data/networks/shift_214_net.Rdata')
net_214 <- graph_from_data_frame(d= shift_214$edges, vertices = shift_214$vertices, directed=F)
plot(net_214)
```

# Shift 217

```{r}
shift_217 <- read_rds('Data/networks/shift_217_net.Rdata')
net_217 <- graph_from_data_frame(d= shift_217$edges, vertices = shift_217$vertices, directed=F)
plot(net_217)

```

# Shift 223

```{r}
shift_223 <- read_rds('Data/networks/shift_223_net.Rdata')
net_223 <- graph_from_data_frame(d= shift_223$edges, vertices = shift_223$vertices, directed=F)
plot(net_223)

```

# Shift 226

```{r}
shift_226 <- read_rds('Data/networks/shift_226_net.Rdata')
net_226 <- graph_from_data_frame(d= shift_226$edges, vertices = shift_226$vertices, directed=F)
plot(net_226)

```

# Shift 230

```{r}
shift_230 <- read_rds('Data/networks/shift_230_net.Rdata')
net_230 <- graph_from_data_frame(d= shift_230$edges, vertices = shift_230$vertices, directed=F)
plot(net_230)
```

# Shift 233

```{r}
shift_233 <- read_rds('Data/networks/shift_233_net.Rdata')
net_233 <- graph_from_data_frame(d= shift_233$edges, vertices = shift_233$vertices, directed=F)
plot(net_233)


```

# Shift 240

```{r}
shift_240 <- read_rds('Data/networks/shift_240_net.Rdata')
net_240 <- graph_from_data_frame(d= shift_240$edges, vertices = shift_240$vertices, directed=F)
plot(net_240)

```

# Shift 243

```{r}
shift_243 <- read_rds('Data/networks/shift_243_net.Rdata')
net_243 <- graph_from_data_frame(d= shift_243$edges, vertices = shift_243$vertices, directed=F)
plot(net_243)

```

# Shift 247

```{r}
shift_247 <- read_rds('Data/networks/shift_247_net.Rdata')
net_247 <- graph_from_data_frame(d= shift_247$edges, vertices = shift_247$vertices, directed=F)
plot(net_247)

```

# Shift 252

```{r}
shift_252 <- read_rds('Data/networks/shift_252_net.Rdata')
net_252 <- graph_from_data_frame(d= shift_252$edges, vertices = shift_252$vertices, directed=F)
plot(net_252)

```

# Shift 258

```{r}
shift_258 <- read_rds('Data/networks/shift_258_net.Rdata')
net_258 <- graph_from_data_frame(d= shift_258$edges, vertices = shift_258$vertices, directed=F)
plot(net_258)

```

# Shift 262

```{r}
shift_262 <- read_rds('Data/networks/shift_262_net.Rdata')
net_262 <- graph_from_data_frame(d= shift_262$edges, vertices = shift_262$vertices, directed=F)
plot(net_262)

```

# Shift 264

```{r}
shift_264 <- read_rds('Data/networks/shift_264_net.Rdata')
net_264 <- graph_from_data_frame(d= shift_264$edges, vertices = shift_264$vertices, directed=F)
plot(net_264)

```

# Shift 266

```{r}
shift_266 <- read_rds('Data/networks/shift_266_net.Rdata')
net_266 <- graph_from_data_frame(d= shift_266$edges, vertices = shift_266$vertices, directed=F)
plot(net_266)

```

# Shift 271

```{r}
shift_271 <- read_rds('Data/networks/shift_271_net.Rdata')
net_271 <- graph_from_data_frame(d= shift_271$edges, vertices = shift_271$vertices, directed=F)
plot(net_271)
```

# Shift 276

```{r}
shift_276 <- read_rds('Data/networks/shift_276_net.Rdata')
net_276 <- graph_from_data_frame(d= shift_276$edges, vertices = shift_276$vertices, directed=F)
plot(net_276)

```

# Shift 278

```{r}
shift_278 <- read_rds('Data/networks/shift_278_net.Rdata')
net_278 <- graph_from_data_frame(d= shift_278$edges, vertices = shift_278$vertices, directed=F)
plot(net_278)

```

# Shift 283

```{r}
shift_283 <- read_rds('Data/networks/shift_283_net.Rdata')
net_283 <- graph_from_data_frame(d= shift_283$edges, vertices = shift_283$vertices, directed=F)
plot(net_283)

```

# Shift 286

```{r}
shift_286 <- read_rds('Data/networks/shift_286_net.Rdata')
net_286 <- graph_from_data_frame(d= shift_286$edges, vertices = shift_286$vertices, directed=F)
plot(net_286)
```

# Shift 289

```{r}
shift_289 <- read_rds('Data/networks/shift_289_net.Rdata')
net_289 <- graph_from_data_frame(d= shift_289$edges, vertices = shift_289$vertices, directed=F)
plot(net_289)

```

# Shift 293

```{r}
shift_293 <- read_rds('Data/networks/shift_293_net.Rdata')
net_293 <- graph_from_data_frame(d= shift_293$edges, vertices = shift_293$vertices, directed=F)
plot(net_293)

```

# Shift 295

```{r}
shift_295 <- read_rds('Data/networks/shift_295_net.Rdata')
net_295 <- graph_from_data_frame(d= shift_295$edges, vertices = shift_295$vertices, directed=F)
plot(net_295)

```

# Shift 298

```{r}
shift_298 <- read_rds('Data/networks/shift_298_net.Rdata')
net_298 <- graph_from_data_frame(d= shift_298$edges, vertices = shift_298$vertices, directed=F)
plot(net_298)
```

# Shift 304

```{r}
shift_304 <- read_rds('Data/networks/shift_304_net.Rdata')
net_304 <- graph_from_data_frame(d= shift_304$edges, vertices = shift_304$vertices, directed=F)
plot(net_304)
```

# Shift 306

```{r}
shift_306 <- read_rds('Data/networks/shift_306_net.Rdata')
net_306 <- graph_from_data_frame(d= shift_306$edges, vertices = shift_306$vertices, directed=F)
plot(net_306)
```

# Shift 308

```{r}
shift_308 <- read_rds('Data/networks/shift_308_net.Rdata')
net_308 <- graph_from_data_frame(d= shift_308$edges, vertices = shift_308$vertices, directed=F)
plot(net_308)
```

# Shift 313

```{r}
shift_313 <- read_rds('Data/networks/shift_313_net.Rdata')
net_313 <- graph_from_data_frame(d= shift_313$edges, vertices = shift_313$vertices, directed=F)
plot(net_313)

```

# Shift 317

```{r}
shift_317 <- read_rds('Data/networks/shift_317_net.Rdata')
net_317 <- graph_from_data_frame(d= shift_317$edges, vertices = shift_317$vertices, directed=F)
plot(net_317)

```

# Shift 320

```{r}
shift_320 <- read_rds('Data/networks/shift_320_net.Rdata')
net_320 <- graph_from_data_frame(d= shift_320$edges, vertices = shift_320$vertices, directed=F)
plot(net_320)

```

# Shift 325

```{r}
shift_325 <- read_rds('Data/networks/shift_325_net.Rdata')
net_325 <- graph_from_data_frame(d= shift_325$edges, vertices = shift_325$vertices, directed=F)
plot(net_325)
```

# Shift 328

```{r}
shift_328 <- read_rds('Data/networks/shift_328_net.Rdata')
net_328 <- graph_from_data_frame(d= shift_328$edges, vertices = shift_328$vertices, directed=F)
plot(net_328)
```



# Shift 331

```{r}
shift_331 <- read_rds('Data/networks/shift_331_net.Rdata')
net_331 <- graph_from_data_frame(d= shift_331$edges, vertices = shift_331$vertices, directed=F)
plot(net_331)
```

# Shift 335

```{r}
shift_335 <- read_rds('Data/networks/shift_335_net.Rdata')
net_335 <- graph_from_data_frame(d= shift_335$edges, vertices = shift_335$vertices, directed=F)
plot(net_335)
```

# Shift 341

```{r}
shift_341 <- read_rds('Data/networks/shift_341_net.Rdata')
net_341 <- graph_from_data_frame(d= shift_341$edges, vertices = shift_341$vertices, directed=F)
plot(net_341)
```

# Shift 346

```{r}
shift_346 <- read_rds('Data/networks/shift_346_net.Rdata')
net_346 <- graph_from_data_frame(d= shift_346$edges, vertices = shift_346$vertices, directed=F)
plot(net_346)
```

# Shift 353

```{r}
shift_353 <- read_rds('Data/networks/shift_353_net.Rdata')
net_353 <- graph_from_data_frame(d= shift_353$edges, vertices = shift_353$vertices, directed=F)
plot(net_353)
```

# Shift 355

```{r}
shift_355 <- read_rds('Data/networks/shift_355_net.Rdata')
net_355 <- graph_from_data_frame(d= shift_355$edges, vertices = shift_355$vertices, directed=F)
plot(net_355)
```

