---
title: "Network Analysis of Clinical Interactions (NACI): Data Cleaning Log"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
# Versioning & Setup  

```{r setup, include=FALSE}
# This is an example setup chunk from the N741 project
knitr::opts_chunk$set(root.dir = "~/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data",
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
# options(na.action = na.warn)??


# Load packages
# library(igraph)
library(haven)
library(tidyverse)
# library(lubridate)
# library(pander)
# library(printr)
# library(forcats)
# library(modelr)

```


Copied from "Data/IPIRC readme file for final code and location-by-second datasets.docx":
Has sid, date, and columns representing every second of the day (a very wide file).
Each row contains the locations (numeric values) for that sid-date.
Recall that sids can repeat over shifts (tags re-used on other patients; staff have permanent tags).

Cpat.zip = all shifts, patient locations per second
Cstaff.zip==all shifts, staff locations per second

*Room locations with square footage are in an Excel file, which links location numbers to location names


```{r data import & transform}

# setwd("~/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data")

# 1a. read "completepat.sas7bdat", 
pt_complete <- 
  read_sas("/Users/tommy-two/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data/completepat.sas7bdat")
# 2a. subset first 10 observations for data transformation code preparation
pt_head <- head(pt_complete) %>% 
  # 3a. Pivot the data.frame from wide to long by placing all column names that start with "floc" into a new column, "time_seconds," and placing respective observations for each "floc" variable into a "location" column
  pivot_longer(cols = starts_with("floc"), names_to = "time_seconds", values_to = "location") %>% 
  # 4a. Remove the prefix "floc" from all observations now under "time_seconds"
  mutate(time_seconds = as.integer(str_replace(time_seconds, "floc", "")), 
         shift_num_ampm = str_trim(shift_num_ampm),
         shift_num = as.integer(str_extract(shift_num_ampm, "[:digit:]+")),
         am_pm = str_extract(shift_num_ampm, "am|pm")) # %>% 
  # 5. Filter out all rows for which no location was recorded
  # filter(!is.na(location))
glimpse(pt_head)
# ----
# 1b. read "completestaff.sas7bdat"
staff_c <- 
  read_sas("/Users/tommy-two/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data/completestaff.sas7bdat")


cc_popn <- read_sas("/Users/tommy-two/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data/Data_Files/chief_complaints_popn.sas7bdat")

# edges <- haven::read_sas("tommy_edges_2009.sas7bdat") 
# read in the node traits: nodes
# nodes <- haven::read_sas("tommy_nodes_2009.sas7bdat")
# read in patient characteristics data: patients
# patients <- haven::read_sas("tommyflynn_patient_info_2009.sas7bdat")

```

```{r Date Fix, eval=FALSE}
# There is an issue with dates; each dataset has a different set of dates; figure out what dates to use with:
PTS_d8s <- unique(patients$d8) 
Nodes_d8s <-  unique(nodes$d8)
Edges_D8s <-  unique(edges$D8)
alldates <- c(PTS_d8s, Nodes_d8s , Edges_D8s)
# length(alldates) answer= 131
alldates <- unique(alldates) # There are 49 total unique dates across all three tbls, but only 35 in edges!
```

```{r Chief Complaint Levels, eval=FALSE}
# Chief complaint is a string with more than 300 variables (mostly referring to the same short-list of complaints (i.e. many duplicates)). Re-level with fct_collapse() in a mutate() call.

patients_cc <- patients %>% mutate(CC_group = fct_collapse(Chief_Complaint,
                                            Pain_Abd = c("Pain, Abd General", "Abdominal pain",
                                                         "Abdominal pain - pregnant", "Abdominal problem",
                                                         "Pain, Abd LLQ", "Nausea/Vomiting with Abd Pain",
                                                         "Flank Pain", "Pain, Abd RLQ", "Pain, Epigastric",
                                                         "Pain, Flank", "Preg w/ Abd Pain", "Pain, Abd RUQ",
                                                         "Flank pain", "Pain, Abd LUQ"),
                                            Pain_Chest = c("Chest Pain", "Chest pain", "Pain, Chest", 
                                                           "Chest Pressure", "Chest Wall Pain", 
                                                           "Heartburn or indigestion",
                                                           "Pain, Chest/Ribs", "Pain, Rib(s)"),
                                            Resp_ARDS = c("Dyspnea", "Difficulty Breathing", "Asthma", 
                                                          "Respiratory problem", "Wheezing", 
                                                          "Exposure, Chemical, Inhalation"),
                                            Pain_MSK = c("Back Pain", "Pain, Knee", "Pain, Leg", 
                                                         "Pain, Ankle", 
                                                         "Pain, Arm", "Pain, Back", "Pain, Foot", 
                                                         "Pain, Hip", 
                                                         "Pain, Other", "Body aches (general)", 
                                                         "Knee pain - swelling", 
                                                         "Pain, Shoulder", "Pain, Neck", 
                                                         "Back pain - lumbar", 
                                                         "Leg pain - swelling", "Pain, Hand", 
                                                         "Arm pain - swelling",
                                                         "Foot pain - swelling", "Pain, Low Back", 
                                                         "Pain, Wrist", 
                                                         "Shoulder pain - swelling", 
                                                         "Ankle pain - swelling",
                                                         "Elbow pain - swelling", 
                                                         "Hip pain - swelling", "Neck pain",
                                                         "Pain, Breast", "Pain, Buttocks", "Pain, Finger", 
                                                         "Pain, Forearm",
                                                         "Pain, Joints, Multiple", "Pain, Muscular General",
                                                         "Pain, Toe",
                                                         "Back pain - thoracic", "Facial pain", 
                                                         "Hand pain - swelling", 
                                                         "Lower leg pain - swelling", "Pain, Face", 
                                                         "Pain, Heel",
                                                         "Multiple musculoskeletal pain/swelling", 
                                                         "Pain, Generalized", 
                                                         "Rib/Trunk pain - swelling", 
                                                         "Wrist pain - swelling", 
                                                         "Finger(s) pain - swelling", "Preg w/ Back Pain"),
                                            Neuro_HA = c("Headache", "Headache, Migraine Hx", 
                                                         "Headache - recurrent",
                                                         "Headache, Cluster Hx", 
                                                         "Headache, New, No Prior Hx", 
                                                         "Headache, Post Traumatic", "Pain, Head"),
                                            Resp_Inf = c("Flu Symptoms", "Cough", "Sore Throat", "URI Sxs", 
                                                         "Cough, Productive", "Throat pain", 
                                                         "Cough, Cold, Congestion",
                                                         "Cold Symptoms", "Congestion", 
                                                         "Upper respiratory infection",
                                                         "Cough, Non Productive"),
                                            Other = c("Multiple Complaints", "AICD discharge", 
                                                      "Medical Problem - Minor", "Other Complaint", 
                                                      "Sweats", "Medical Problem - Minor", "Hiccoughs",
                                                      "Choking", "Cyst", "Gout"),
                                            Neuro_AMS = c("Dizzy", "Syncope", "Altered Mental Status", 
                                                          "Dizziness", 
                                                          "Altered mental status", "Seizure, Epilepsy Hx",
                                                          "Fatigue",
                                                          "Seizure, Non Specific", "Neurological problem",
                                                          "Near Syncope",
                                                          "Seizure, New Onset", "Difficulty Speaking", 
                                                          "Confusion", "Fainting", "Seizure", 
                                                          "Seizure - prior history", 
                                                          "Vertigo", "Vision Loss or Disturbance"),
                                            GI_General = c("Vomiting", "Bleeding, Rectal",
                                                           "Nausea/Vomiting/Diarrhea", 
                                                         "Nausea/Vomiting", "Diarrhea", "Nausea", 
                                                         "GI bleed",
                                                         "Constipation", "Nausea/Vomiting/ Abd Pain",
                                                         "Dehydration", 
                                                         "Rectal pain", "FB, Rectum", "Hemorrhoids",
                                                         "Hernia, Abd",
                                                         "Hernia, Other", "Loss of Appetite", 
                                                         "Failure to Thrive",
                                                         "Pain, Hemorroids", "Preg w/ Epigastric Pain", 
                                                         "Preg w/ Nausea,Vomiting &lt; 20 wks",
                                                         "Regurgitation", 
                                                         "Vomiting - pregnant"),
                                            GU = c("Bleeding, Vaginal", "Vaginal Discharge", 
                                                   "Urinary Retention", 
                                                   "Urinary Burning", "Pain, Pelvic", 
                                                   "Pelvic pain &lt; 20 weeks pregnant",
                                                   "Preg, Vag Bleed, &lt; 20 wks", "Dysuria", "Hematuria",
                                                   "Pelvic pain",
                                                   "STD exposure", "Vaginal bleed", 
                                                   "Vaginal discharge/Itching", 
                                                   "Dysuria/Frequency", "Genitourinary problem", 
                                                   "Preg w/ Pelvic Pain",
                                                   "Preg w/ painful vaginal bleeding", "STD Exposure",
                                                   "Vaginal pain",
                                                   "Vaginal bleed &lt; 20 weeks pregnant", "FB, Vagina",
                                                   "Groin pain", 
                                                   "Incontinent, Urine", "Pain, Groin", "Penile Discharge",
                                                   "Preg w/ Vag D/C", "Preg w/ Vag Pain", 
                                                   "Preg, Poss Spont Abortion",
                                                   "Preg, Possible", "Testicular pain", "Urinary Frequency",
                                                   "Urinary Hesitancy", "STD, Possible"),
                                            Weakness_Numbness = c("Weakness, Generalized", 
                                                          "Weakness, All Extremities", 
                                                         "Weakness or fatigue", "Focal motor weakness",
                                                         "Weakness, One Side Body", "Numbness, Arm(s)", 
                                                         "Numbness, One Side Body", "Numbness, Face", 
                                                         "Numbness, Hand(s)/Finger(s)", "Numbness, Other",
                                                         "Weakness, Face",
                                                         "Weakness, Lower Ext", "Weakness, Other", 
                                                         "Weakness, Upper Ext"),
                                            MVC = c("Motor Vehicle Crash", "MVC - Minor", 
                                                    "Auto-Pedest Crash"),
                                            Fall = c("Fall", "Fall, ground level"),
                                            Swelling = c("Swelling, Leg", "Swelling, Face", 
                                                         "Edema, Lower Ext",
                                                         "Swelling, Hand", "Swelling, Knee", 
                                                         "Swelling, Neck", "Edema, Other",
                                                         "Swelling, Abd", "Swelling, Arm", "Swelling, Eye", 
                                                         "Swelling, Foot",
                                                         "Swollen Glands Other", "Leg edema", 
                                                         "Swelling, Airway", "Abnormal diagnostic test",
                                                         "Swelling, General", "Swelling, Ankle"),
                                            Glycemias = c("Hyperglycemia", "Hypoglycemia", 
                                                          "Diabetes Mellitus, IDDM"),
                                            Heme_HgSS = c("Sickle Cell crisis", "Sickle Cell, Pain", 
                                                          "Sickle Cell, Joint Pain",
                                                          "Sickle Cell, Chest Pain", "Sickle Cell, Other"),
                                            Allergic_Rx = c("Allergic Reaction", 
                                                            "Allergic reaction - minor", 
                                                           "Allergic reaction - major", 
                                                           "Medication Reaction"),
                                            Follow_Up = c("Surgical Problem Re-evaluation", 
                                                          "Medical Clearance",
                                                          "Infection, Incision", "Lab draw", 
                                                          "Medication refill", 
                                                          "Suture Removal", "Abnormal diagnostic test",
                                                          "Surgical problem re-evaluation",
                                                          "Wound, Dehiscence", "Wound, Re-eval", 
                                                          "Supplies Only", "Lab or Test Abnormality",
                                                          "Medical problem re-evaluation", 
                                                          "Disloc Hip, Prior Replace"),
                                            Chronic = c( "Catheter check or insertion", 
                                                         "Dialysis shunt problem", 
                                                         "Dialysis Access Problem", "Cancer", 
                                                         "Pacemaker Problem",
                                                         "Dialysis Prob, Hemodialysis Access", 
                                                         "Dialysis Prob, Hemodialysis Graft", 
                                                         "GI tube evaluation", 
                                                         "ICD Related Problem", 
                                                         "Indwelling line problem",
                                                         "PEG Tube Problem", "PICC Line Problem"),
                                            Behavioral = c("Suicidal", "Anxiety/Anxious", 
                                                           "Alcohol Intoxication Possible",
                                                           "Alcohol Related Problem", "Depression", 
                                                           "Psych, Anxiety/Anxious",
                                                           "Suicidal ideation", "Anxiety", 
                                                           "Detox Requested", "Suicide Attempt", 
                                                           "Suicidal Ideation", "Withdrawal, Alcohol",
                                                           "Psychiatric problem", 
                                                           "Psych, Behavior Disorder"),
                                            Injury = c("Assault, Physical", "Sexual assault", "Lac, Finger",
                                                       "Bleeding, Other", "Lac, Leg", "Inj, Finger(s)",
                                                       "Lac, Hand", "Hip injury - minor", "Inj, Arm",
                                                       "Inj, Hand(s)", "Inj, Leg", "Inj, Head", 
                                                       "Inj, Wrist", "Lac, Arm", "Assault, Sexual",
                                                       "Lac, Face", "Lac, Forehead", "Lac, Scalp", 
                                                       "Facial injury - minor", "Extremity injury - minor",
                                                       "Inj, Ankle", "Inj, Hip", "Lac, Chin",
                                                       "Knee injury - minor", "Lac, Eyelid", "Lac, Foot",
                                                       "Lac, Lip", "Laceration - facial", 
                                                       "Laceration - finger", "Laceration - scalp",
                                                       "Laceration - facial", "Laceration - finger",
                                                       "Laceration - scalp",
                                                       "Puncture Wound", "Shoulder injury - minor", 
                                                       "Wrist injury - minor",
                                                       "Hand injury - major", "Arm injury - minor",
                                                       "Abrasion minor", 
                                                       "Abrasion, Abd", "Bite, Insect", "Bite, Dog", 
                                                       "Bite, Human", 
                                                       "Bite, Spider", "Bite, Animal", 
                                                       "Bite, Unknown", "Sting, Insect"),
                                            HEENT_oral = c("Eye(s), Complaint, Other", "Pain, Ear",
                                                           "Earache", "Nosebleed", 
                                                      "ENT problem", "Pain, Eye", "Pain, Dental", 
                                                      "Pain, Oral", "Eye pain",
                                                      "Swallowing Problem", "Ear Discharge", "Epistaxis", 
                                                      "Ear pain", 
                                                      "Eye(s), Swollen", "Hemoptysis", "Dental pain", 
                                                      "Difficulty swallowing",
                                                      "Ear Wax", "Eye foreign body", "Eye(s), Irritation", 
                                                      "FB, Throat", 
                                                      "Nasal drainage", "Tongue swelling"),
                                            CV_Sxs = c("Hypertension", "Hypotension", "Tachycardia", 
                                                       "Heartbeat, Palpitations",
                                                       "Palpitations", "Heartbeat, Irregular", 
                                                       "Bradycardia", 
                                                       "Heartbeat, Racing"),
                                            Fever = c("Fever, Pediatric", "Fever, Adult", "Fever",
                                                       "Chills", "Fever, Immunocompromised"),
                                           Skin_Wnd_Absc = c("Rash", "Itching", "Sores, Skin", 
                                                     "Burn, Thermal","Abscess - complicated", 
                                                     "Abscess - simple", "Abscess Skin/Soft Tissue", 
                                                     "Infection, Skin/Soft Tissue, Other", "Jaundice",
                                                     "Burn, Chemical", "Infection, Hand/Finger",
                                                     "Infection, Other", "Poison Ivy/Poison Oak", 
                                                     "Rash, Possible Allergy", 
                                                     "Wound infection (complicated)",
                                                     "Skin lesion bleed", "Skin problem", "Leg ulcers", 
                                                     "Ulcer, Lower Extremity", 
                                                     "Infection, Rectal/Perirectal" ),
                                            Emergent = c("Medical Problem - Major", "Wound, Gunshot", 
                                                       "Cardiac and/or respiratory arrest")
                                            
))
```

```{r chunk-tidy?, eval=FALSE}
# names(nodes)
# Now clean the patients data to include: acuity recorded, race black or white, date in edges, chief complaint, MinutesInED, 
patients35 <- patients_cc %>% filter(Acuity != "Not Recorded", !is.na(Acuity), Race != "2") %>% 
  mutate(Race = as.factor(Race), Acuity = as.factor(Acuity), Sex = as.factor(Sex)) %>%
  select(sid, d8, Acuity, Race, AGE, Sex, Chief_Complaint, CC_group, MinutesInED, ED_ARRIVAL, 
         ed_departure, Arr_Mode, ED_Disposition, numshift)
# head(patients)
# Now tidy Edges, select only the needed variables (sidi, sidj, participant type i-j, date, am, pm, 
# weight); remove all edges between two patients

edges35 <- edges %>% filter(!is.na(i_participant_type), !is.na(j_participant_type), !is.na(i), !is.na(j),
                            (i_participant_type  != "_PAT" | j_participant_type != "_PAT" )) %>%
  mutate(i = i_participant_type, j = j_participant_type, am = (shiftampm == 1), 
         pm = (shiftampm == 2)) %>%
  select(sidi, sidj, i, j, edgeweight, am, pm, D8, idi, idj, numshift)

# Alright; lets get to the nodes.. need to have the ID come first...
nodes35 <- nodes %>% filter(Race != "2") %>%
  mutate(ID = as.character(ID), i = as.character(i), AGE = as.integer(AGE), Acuity = as.factor(Acuity), 
                            Race = as.factor(Race), Sex = as.factor(Sex)) %>%
  select(sid, i, d8, ID, participant_type, AGE, Race, Sex, Acuity, Arr_Mode, shift_num,
         MinutesInED,  duration_observed, shift_ampm, ED_Disposition, BLACKyn, participationrate)
```

```{r chunk-participant attributes, eval=FALSE}
# patient demographics
# Race stats
racedf <- data_frame("Race" = c("Black", "Hispanic", "Other", "White"), "Count" = summary(patients35$Race))
race_table <- knitr::kable(racedf, align = "l", caption = "Patient Race",
             row.names = FALSE)
race_plot <- ggplot(patients35, aes(x = AGE)) + geom_histogram() + 
  scale_x_continuous("Patient Age Distribution") + 
  labs(title = "Distribution of Patient Age")
# Acuity stats
acuitydf <- data_frame("Acuity Level" = c("Immediate (1)", "Emergent (2)", "Urgent (3)", "Stable (4)", "Non Urgent (5)"),
                       "Count" = summary(patients35$Acuity))
acuity_table <- knitr::kable(acuitydf, align = "l", caption = "Patient Acuity",
             row.names = FALSE)
acuity_ploot <- ggplot(patients35, aes(x = Acuity, fill = Acuity)) + geom_bar(position = "dodge") + 
  scale_x_discrete("Patient Acuity") + labs(title = "Distribution of Patient Acuity")
# Gender Stats
gender <- data_frame("Sex" = c("Female", "Male"), "Count" = summary(patients35$Sex))
gender_table <- knitr::kable(gender, align = "l", caption = "Patient Gender",
             row.names = FALSE)
# Age stats
mean_age <- patients35 %>% summarise(Age_mean = mean(AGE, na.rm=TRUE), sd = sd(AGE, na.rm=TRUE))
mean_age <- knitr::kable(mean_age, align = "l", caption = "Patient Age",
             row.names = FALSE)
age_distribution <- ggplot(patients35, aes(x = AGE)) + geom_histogram(fill = "orange", alpha = .5) + 
  scale_x_continuous() + 
  labs(title = "Distribution of Patient Age", y = "Count", x = "Patient Age")
# now let's see if this works... finally
participants <- nodes35 %>% summarise(n = n_distinct(ID), shifts = n_distinct(shift_num), meanparts = n/35,
                                  participation = mean(participationrate), ties = length(edges35$sidi), 
                                 tieshifts = ties/35)
headers <- c("Participants (n)", "Shifts", "Participants/Shift", "Participation Rate (mean%)", "Total Ties",
            "Ties/Shift")
participants <- knitr::kable(participants, align = "c", caption = "Overall Participation", 
             col.names = headers, row.names = FALSE)
# Chief Complaints
CC_graph <- ggplot(patients35, aes(x = Chief_Complaint, fct_reorder(CC_group, MinutesInED), 
                                   color = Acuity)) + geom_count() + theme(legend.position = "right")
```

```{r chunk-unused plots, eval=FALSE}
ccgrou_plot <- ggplot(patients35, aes(AGE, fct_reorder(CC_group, MinutesInED), color = Acuity)) +
         geom_count(position = "jitter", alpha = 0.6, shape = 1, size = 4) +
         scale_x_continuous("Patient Age (yrs)", limits = c(1, 100)) + 
         scale_y_discrete(name = "Chief Complaint")
disposition_plot <- ggplot(patients35, aes(ED_Disposition)) + geom_bar() + 
  scale_x_discrete("ED Disposition")
```

```{r chunk-shift 35, eval=FALSE}
# Edges by shift... this could have been easier; but I spent too much time working my way through the data already; no time to learn how to write a function or loop... :(

# Edges by shift... this could have been easier; but I spent too much time working my way through the data already; no time to learn how to write a function or loop... :(
Eshift1 <- edges35 %>% filter(D8 == Edges_D8s[1])
Eshift2 <- edges35 %>% filter(D8 == Edges_D8s[2])
Eshift3 <- edges35 %>% filter(D8 == Edges_D8s[3])
Eshift4 <- edges35 %>% filter(D8 == Edges_D8s[4])
Eshift5 <- edges35 %>% filter(D8 == Edges_D8s[5])
Eshift6 <- edges35 %>% filter(D8 == Edges_D8s[6])
Eshift7 <- edges35 %>% filter(D8 == Edges_D8s[7])
Eshift8 <- edges35 %>% filter(D8 == Edges_D8s[8])
Eshift9 <- edges35 %>% filter(D8 == Edges_D8s[9])
Eshift10 <- edges35 %>% filter(D8 == Edges_D8s[10])
Eshift11 <- edges35 %>% filter(D8 == Edges_D8s[11])
Eshift12 <- edges35 %>% filter(D8 == Edges_D8s[12])
Eshift13 <- edges35 %>% filter(D8 == Edges_D8s[13])
Eshift14 <- edges35 %>% filter(D8 == Edges_D8s[14])
Eshift15 <- edges35 %>% filter(D8 == Edges_D8s[15])
Eshift16 <- edges35 %>% filter(D8 == Edges_D8s[16])
Eshift17 <- edges35 %>% filter(D8 == Edges_D8s[17])
Eshift18 <- edges35 %>% filter(D8 == Edges_D8s[18])
Eshift19 <- edges35 %>% filter(D8 == Edges_D8s[19])
Eshift20 <- edges35 %>% filter(D8 == Edges_D8s[20])
Eshift21 <- edges35 %>% filter(D8 == Edges_D8s[21])
Eshift22 <- edges35 %>% filter(D8 == Edges_D8s[22])
Eshift23 <- edges35 %>% filter(D8 == Edges_D8s[23])
Eshift24 <- edges35 %>% filter(D8 == Edges_D8s[24])
Eshift25 <- edges35 %>% filter(D8 == Edges_D8s[25])
Eshift26 <- edges35 %>% filter(D8 == Edges_D8s[26])
Eshift27 <- edges35 %>% filter(D8 == Edges_D8s[27])
Eshift28 <- edges35 %>% filter(D8 == Edges_D8s[28])
Eshift29 <- edges35 %>% filter(D8 == Edges_D8s[29])
Eshift30 <- edges35 %>% filter(D8 == Edges_D8s[30])
Eshift31 <- edges35 %>% filter(D8 == Edges_D8s[31])
Eshift32 <- edges35 %>% filter(D8 == Edges_D8s[32])
Eshift33 <- edges35 %>% filter(D8 == Edges_D8s[33])
Eshift34 <- edges35 %>% filter(D8 == Edges_D8s[34])
Eshift35 <- edges35 %>% filter(D8 == Edges_D8s[35])
# Nodes by shift
Vshift1 <- nodes35 %>% filter(d8 == Edges_D8s[1])
Vshift2 <- nodes35 %>% filter(d8 == Edges_D8s[2])
Vshift3 <- nodes35 %>% filter(d8 == Edges_D8s[3])
Vshift4 <- nodes35 %>% filter(d8 == Edges_D8s[4])
Vshift5 <- nodes35 %>% filter(d8 == Edges_D8s[5])
Vshift6 <- nodes35 %>% filter(d8 == Edges_D8s[6])
Vshift7 <- nodes35 %>% filter(d8 == Edges_D8s[7])
Vshift8 <- nodes35 %>% filter(d8 == Edges_D8s[8])
Vshift9 <- nodes35 %>% filter(d8 == Edges_D8s[9])
Vshift10 <- nodes35 %>% filter(d8 == Edges_D8s[10])
Vshift11 <- nodes35 %>% filter(d8 == Edges_D8s[11])
Vshift12 <- nodes35 %>% filter(d8 == Edges_D8s[12])
Vshift13 <- nodes35 %>% filter(d8 == Edges_D8s[13])
Vshift14 <- nodes35 %>% filter(d8 == Edges_D8s[14])
Vshift15 <- nodes35 %>% filter(d8 == Edges_D8s[15])
Vshift16 <- nodes35 %>% filter(d8 == Edges_D8s[16])
Vshift17 <- nodes35 %>% filter(d8 == Edges_D8s[17])
Vshift18 <- nodes35 %>% filter(d8 == Edges_D8s[18])
Vshift19 <- nodes35 %>% filter(d8 == Edges_D8s[19])
Vshift20 <- nodes35 %>% filter(d8 == Edges_D8s[20])
Vshift21 <- nodes35 %>% filter(d8 == Edges_D8s[21])
Vshift22 <- nodes35 %>% filter(d8 == Edges_D8s[22])
Vshift23 <- nodes35 %>% filter(d8 == Edges_D8s[23])
Vshift24 <- nodes35 %>% filter(d8 == Edges_D8s[24])
Vshift25 <- nodes35 %>% filter(d8 == Edges_D8s[25])
Vshift26 <- nodes35 %>% filter(d8 == Edges_D8s[26])
Vshift27 <- nodes35 %>% filter(d8 == Edges_D8s[27])
Vshift28 <- nodes35 %>% filter(d8 == Edges_D8s[28])
Vshift29 <- nodes35 %>% filter(d8 == Edges_D8s[29])
Vshift30 <- nodes35 %>% filter(d8 == Edges_D8s[30])
Vshift31 <- nodes35 %>% filter(d8 == Edges_D8s[31])
Vshift33 <- nodes35 %>% filter(d8 == Edges_D8s[33])
Vshift34 <- nodes35 %>% filter(d8 == Edges_D8s[34])
Vshift35 <- nodes35 %>% filter(d8 == Edges_D8s[35])
```

```{r chunk-relationships, eval=FALSE}
# Plot acuity by different degree measures
acuityEIGEN <- nodes %>% filter(participant_type == "_PAT", Acuity != "Not Recorded") %>%
  ggplot(aes(x = Acuity, y = Eigenvector_Centrality, color = Acuity)) + 
  geom_boxplot(shape = 1)
acuitybyWTDEG <- nodes %>% filter(participant_type == "_PAT", Acuity != "Not Recorded") %>%
  ggplot(aes(x = LOG_WTDEGREE, y = Acuity, color = Acuity)) + 
  geom_boxplot(shape = 1)

```

```{r chunk-shift 10 a, eval=FALSE}
# head(Eshift1)
# head(Vshift1)
# similar <- Eshift1$sidi %in% Vshift1$sid
# table(similar)
# similar2 <- Eshift1$idi %in% Vshift1$ID
# table(similar2)
Vshift10 %>% mutate(Acuity = as.factor(Acuity)) -> Vshift10
shift10 <- graph_from_data_frame(d = Eshift10, directed = FALSE, vertices = Vshift10)
V(shift10)$label <- Vshift10$participant_type
V(shift10)$gender <- Vshift10$Sex
E(shift10)$weight <- Eshift10$edgeweight
V(shift10)$acuity <- Vshift10$Acuity
# set vertex attributes(shift10)
# vertex color
colrs <- c("gray50", "blue", "gold", "tomato", "white")
V(shift10)$color <- colrs[V(shift10)$acuity]
# Vertex size [add when centrality is calculated]
# V(shift10)$size <- V(shift10)$MinutesInED
# EDge width
E(shift10)$width <- E(shift10)$weight*2
# Set edge color (optional)
E(shift10)$edge.color <- "black"
plot(shift10, vertex.label.dist=2)
legend(x=-1.5, y=-1.1, c("Immediate(1)", "Emergent(2)", "Urgent(3)", "Stable(4)", "Non Urgent(5)"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
```

```{r chunk-shift10 b, eval=FALSE}
plot(shift10, vertex.shape="none", vertex.label=V(shift10)$label, 
     vertex.label.font=2, vertex.label.color="black", vertex.label.cex=.7, edge.color="gray85")
```

```{r chunk-shift10 c, eval=FALSE}
l <- layout_in_circle(shift10)

plot(shift10, layout=l, vertex.label.dist = 2)
```

```{r chunk-Shift10 d, eval=FALSE}

fr <- layout_with_fr(shift10)
fr <- norm_coords(fr, ymin=-1, ymax=1, xmin=-1, xmax=1)
par(mfrow=c(2,2), mar=c(0,0,0,0))
plot(shift10, rescale=F, layout=fr*0.4)
plot(shift10, rescale=F, layout=fr*0.6)
plot(shift10, rescale=F, layout=fr*0.8)
plot(shift10, rescale=F, layout=fr*1.0)



# plot(shift10, vertex.label.dist = 2, layout = fr)
# legend(x=-1.5, y=-1.1, c("Imediate", "Emergent", "Urgent", "Stable", "Non Urgent"), pch=21,
       # col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
# Eshift1$i=="_PAT"
```

```{r chunk-shift 23 a, eval=FALSE}
Vshift23 %>% mutate(Acuity = as.factor(Acuity)) ->Vshift23
shift23 <- graph_from_data_frame(d = Eshift23, directed = FALSE, vertices = Vshift23)
V(shift23)$label <- Vshift23$participant_type
V(shift23)$gender <- Vshift23$Sex
E(shift23)$weight <- Eshift23$edgeweight
V(shift23)$acuity <- Vshift23$Acuity
# set vertex attributes(shift10)
# vertex color
colrs <- c("gray50", "blue", "gold", "tomato", "white")
V(shift23)$color <- colrs[V(shift23)$acuity]
# Vertex size [add when centrality is calculated]
# V(shift10)$size <- V(shift10)$MinutesInED


# EDge width
E(shift23)$width <- E(shift23)$weight*2

# Set edge color (optional)
E(shift23)$edge.color <- "black"



plot(shift23, vertex.label.dist=2)
legend(x=-1.5, y=-1.1, c("Immediate(1)", "Emergent(2)", "Urgent(3)", "Stable(4)", "Non Urgent(5)"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)


```

```{r chunk-shift23 b, eval=FALSE}
layouts <- grep("^layout_", ls("package:igraph"), value=TRUE)[-1] 

# Remove layouts that do not apply to our graph.

layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree", layouts)]

par(mfrow=c(3,3), mar=c(1,1,1,1))

for (layout in layouts) {

  l <- do.call(layout, list(shift23)) 
  
  plot(shift23, edge.arrow.mode=0, layout=l, main=layout) }
```

```{r chunk-shift13, eval=FALSE}
Vshift13 %>% mutate(Acuity = as.factor(Acuity)) -> Vshift13
shift13 <- graph_from_data_frame(d = Eshift13, directed = FALSE, vertices = Vshift13)
V(shift13)$label <- Vshift13$participant_type
V(shift13)$gender <- Vshift13$Sex
E(shift13)$weight <- Eshift13$edgeweight
V(shift13)$acuity <- Vshift13$Acuity
# set vertex attributes(shift10)
# vertex color
colrs <- c("gray50", "blue", "gold", "tomato", "white")
V(shift13)$color <- colrs[V(shift13)$acuity]
# Vertex size [add when centrality is calculated]
# V(shift10)$size <- V(shift10)$MinutesInED
# EDge width
E(shift13)$width <- E(shift13)$weight*2
# Set edge color (optional)
E(shift13)$edge.color <- "black"
num13 <- plot(shift13, vertex.label.dist=2)
  legend(x=-1.5, y=-1.1, c("Immediate(1)", "Emergent(2)", "Urgent(3)", "Stable(4)", 
                           "Non Urgent(5)"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
```

```{r chunk-i, eval=FALSE}
layouts <- grep("^layout_", ls("package:igraph"), value=TRUE)[-1] 

# Remove layouts that do not apply to our graph.

layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree", layouts)]


par(mfrow=c(3,3), mar=c(1,1,1,1))

for (layout in layouts) {

  l <- do.call(layout, list(shift13)) 
  
  plot(shift13, edge.arrow.mode=0, layout=l, main=layout) }
```
