---
title: 'Network Analysis of Clinical Interactions (NACI): Data Cleaning Log'
author: 'Tommy Flynn'
output: 
    prettydoc::html_pretty:
      theme: leonids
      toc: true
      number_sections: true
      df_print: paged
---  

# Setup  

## Set chunk options & load package libraries  
```{r Setup}
# set options
# This is an example setup chunk from the N741 project
knitr::opts_chunk$set(root.dir = "~/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data",
                      results = "asis",
                      echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      tidy = TRUE,
                      tidy.opts = list(width.cutoff = 55))
# options(na.action = na.warn)??

```

```{r Libraries}
# Load packages
library(igraph) # package for working with and visualizing network analysis objectve
library(haven) # package for importing SAS data files (i.e., ".sas7bdat")
library(tidyverse) # packages for data import, cleaning, transformation, and analysis
library(gt) # package for creating and formating latex tables
library(lubridate) # package for working with date data
library(knitr)
# library(pander) # ????
# library(printr) # ????
# library(forcats) # package for making and working with factors
# library(modelr) # package for statistical modeling in r
library(readxl)
library(readr)
library(stringr)
library(labelled)
library(details)
library(kableExtra)
```

## Data direcotry    
Make a path object for the data directory and print a list of all data files

```{r Data paths}
data_path <- paste(getwd(), "Data", sep = "/")

kable(tibble(list.files(data_path)))
```

  
  
# Data Import & Cleaning  

## RFID badge location and space data  

### Patient Location: `pt_complete`  
RFID badge location room number for every second of every shift, patients only.  

```{r ReadData pt_complete}
# 1a. read "completepat.sas7bdat",  
pt_complete <- 
  read_sas(paste(data_path, "completepat.sas7bdat", sep = "/"))
```
Large data.frame, using the first six observations to code for data cleaning. The table is extremely wide (>4300 columns), I used `pivot_longer()` to reshape it by collapsing all location-by-second columns into two columns, names to `seconds` and values to `location`. This process causes there to be __many repeated SIDs__.  

```{r CleanData pt_complete}
# 2a. subset first 10 observations for data transformation code preparation
pt_head <- head(pt_complete) %>% 
# 3a. Pivot the data.frame from wide to long by placing all column names that start with "floc" into a new column, "seconds," and placing respective observations for each "floc" variable into a "location_num" column
  pivot_longer(cols = starts_with("floc"), names_to = "seconds", values_to = "location_num") %>% 
# 4a. Remove the prefix "floc" from `time_seconds` and keep the digits as `seconds`
  mutate(seconds = as.integer(str_replace(seconds, "floc", "")), 
         shift_num_ampm = str_trim(shift_num_ampm),
         shift_num = as.integer(str_extract(shift_num_ampm, "[:digit:]+")),
         am_pm = str_extract(shift_num_ampm, "am|pm"),
         date = make_date(year = year, month = mon, day = day)) # %>% 
# 5. Filter out all rows for which no location was recorded
# filter(!is.na(location)) %>% 

# View data frame structure
# glimpse(pt_head)
kable(head(pt_head)) 

```

"Cpat.zip" and "Cstaff.zip" contain "completepat.sas7bdat" and "completestaff.sas7bdat," respectively, that contain location information for patients and staff from all observed shifts, respectively.  
* Both completeXXX.sas7bdat tables have columns for every second of the day, named with the prefix "floc" followed by the second  
* Each row contains the locations (numeric values) for the respective SID and date combination  
  +   Some patient SIDs repeat in the data because RFID tags were used by more than one patient per shift  
  +   Staff had permanent tags, so SID numbers were not duplicated  
* Room locations with square footage are in an Excel file, which links location numbers to location names  
* Need to figure out how to filter out redundancy without cutting data.  

### Staff location: `staff_complete`  
```{r ReadData staff_complete}
# ---- `staff_complete
# 1b. read "completestaff.sas7bdat"
staff_complete <- 
  read_sas(paste(data_path, "completestaff.sas7bdat", sep = "/"))
```

```{r CleanData staff_complete}
# 2a. subset first 10 observations for data transformation code preparation
staff_head <- head(staff_complete) %>% 
  # 3a. Pivot the data.frame from wide to long by placing all column names that start with "floc" into a new column, "seconds," and placing respective observations for each "floc" variable into a "location_num" column
  pivot_longer(cols = starts_with("floc"), names_to = "seconds", values_to = "location_num") %>% 
  # 4a. Remove the prefix "floc" from `time_seconds` and keep the digits as `seconds`
  mutate(seconds = as.integer(str_replace(seconds, "floc", "")), 
         shift_num_ampm = str_trim(shift_num_ampm),
         shift_num = as.integer(str_extract(shift_num_ampm, "[:digit:]+")),
         am_pm = str_extract(shift_num_ampm, "am|pm"),
         date = make_date(year = year, month = mon, day = day)) # %>% 
  # 5. Filter out all rows for which no location was recorded
  # filter(!is.na(location)) %>% 
# View data frame structure
# str(staff_head)
# glimpse(staff_head)
kable(head(staff_head))
```

### List of RFID badge encounters (i.e., clinical interactions) as vertices: `edge_list`  
Read & print data from "allshifts_edges.sas7bdat" and "edges2.sas7bdat."  
```{r ReadData edge_list}
# ---- 'edge_list` ----
# 1c. read "allshifts_edges.sas7bdat"
edge_list <- read_sas(paste(data_path, "allshifts_edges.sas7bdat", sep = "/"))
edge_list2 <- read_sas(paste(data_path, "edges2.sas7bdat", sep = "/"))
```

```{r CleanData edge_list}
# Print the first 6 observations of edge_list
kable(head(edge_list))
# Print out the variable labels for all columns of edge_list (object varbles1)
varbles1 <- var_label(edge_list)
paste(names(varbles1), varbles1, sep = ": ")
# Print the first 6 observations of edge_list2
kable(head(edge_list2))
# Print out the variable labels for all columns of edge_list2 (object varbles2)
varbles2 <- var_label(edge_list2)
paste(names(varbles2), varbles2, sep = ": ")
```
  
## ED Geography:  

List of files in "Data/Data_Reference/Geography":  
```{r ListFiles Geography}
# Create a list of all items in the current working directory
list.files(paste(data_path, "/Data_Reference/Geography"))
# Print directory file list
```


#### Room categories/locations: `room_categories`  

```{r ReadData room_categories}  
room_categories <- read_xlsx(paste(data_path, "Data_Reference/Geography/room categories.xlsx"))
room_categories2 <- read_xlsx(paste(data_path, "Data_Reference/Geography/room categories2.xlsx"))
room_categories20131204 <- read_xlsx(paste(data_path, "Data_Reference/Geography/room categories2 20131204.xlsx"))
room_categories20131204_area <- read_xlsx(paste(data_path, "Data_Reference/Geography/room categories2 20131204 area.xlsx"))
room_location <- read_xlsx(paste(data_path, "Data_Reference/Geography/Room Locations and Square Footage - with Correction.xlsx"))
```


```{r CleanData room_categories}  
geographic_lst <- list(room_categories, room_categories2, room_categories20131204, room_categories20131204_area, room_location)
# Print the first 6 rows form each room category/location file iteratively with `purrr::map()`
map(geographic_lst, c(head, kable))
```
     
   
## Patient attributes  

### Patient ID to RFID badge SID key: `id_sid`  

```{r }
pt_list <- read_sas(paste(data_path, "patient_list_good_2_last.sas7bdat", sep = "/"))
kable(head(pt_list))
```

Read & print data from "id_sid_matchup.sas7bdat" & "id_sid_matchup.sas7bdat."  
```{r ReadData id_sid}
# 1d. read "id_sid_matchup.sas7bdat" into id_sid and "id_sid_matchup2.sas7bdat" into id_sid2
id_sid <- read_sas(paste(data_path, "Data_Files/id_sid_matchup.sas7bdat"))
id_sid2 <- read_sas(paste(data_path, "Data_Files/id_sid_matchup2.sas7bdat"))
```

```{r CleanData id_sid}
# Print the first 6 rows of id_sid
kable(head(id_sid))
# Print the first 6 rows of id_sid2
kable(head(id_sid2))
# Print variable labels
idsid_varbles <- var_label(id_sid)
idsid2_varbles <- var_label(id_sid2)

idsid_labeltable <- paste(names(idsid_varbles), idsid_varbles, sep = ": ")
kable(tibble(idsid_labeltable))

idsid2_labeltable <- paste(names(idsid2_varbles), idsid2_varbles, sep = ": ")
kable(tibble(idsid2_labeltable))


```

### Patient Acuity: `pt_acuity`  
Read & print patient acuity data in "ACUITY-patients.xlsx," which is an Excel workbook with 3 sheets.  

```{r ReadData pt_acuity}
# 1e. read "ACUITY-patients.xlsx
pt_acuity <- read_xlsx(paste(data_path, "Data_Files/ACUITY-patients.xlsx"))
# str(pt_acuity)
pt_acuity_s2 <- read_xlsx(paste(data_path, "Data_Files/ACUITY-patients.xlsx"), sheet = 2, range = "A2:H83")
# str(pt_acuity_s2)
pt_acuity_s3 <- read_xlsx(paste(data_path, "Data_Files/ACUITY-patients.xlsx"), sheet = 3, range = "A2:G83")
# str(pt_acuity_s3)  
```
The first sheet lists the number of patients in each ESI acuity level (columns) by shift (rows).  The other two sheets appear to be variations of the first.  
```{r CleanData pt_acuity}
# Patient acuity (Emergency Severity Index; ESI) counts by shift
kable(head(pt_acuity))
# Pivot_wider to view number of patients in each ESI category by shift
pt_acuity %>% 
  group_by(Acuity) %>% 
  count(Shift) %>% 
  pivot_wider(names_from = Acuity, values_from = n) %>% 
  head() %>% kable()
# Print the first 6 rows of the other two sheets in the xlsx file
kable(head(pt_acuity_s2))
kable(head(pt_acuity_s3))
```


### Patient demographics: `pt_attributes`  

```{r ReadData pt_attributes, eval = FALSE}
pt_attributes <- read_sas(paste(data_path, "attributes.sas7bdat", sep = "/"))
```

```{r CleanData pt_attributes}

```


  
# SessionInfo  
```{r SessionInfo}
sessioninfo::session_info()%>%
  details::details(
    summary = 'Current session info',
    open    = FALSE
  )
```

  
