---
title: "NACI Data Prep"
author: "Tommy J. Flynn"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---
# WRANGLING


```{r setup, message=FALSE}
#| code-fold: true 
knitr::opts_chunk$set(error = FALSE,
                      tidy = TRUE,
                      # tidy.opts = list(width.cutoff = 60),
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(readxl)
library(readr)
library(haven)
library(knitr)
library(kableExtra)
library(lubridate)
library(labelled)
library(gtsummary)
library(data.table)
```

```{r directories}
#| code-fold: true 
# Working directory saved as `wd` and data path as `data_path`  
wd <- getwd()
data_path <- paste(wd, "Data", sep = "/")
```

## Identification numbers 

## `id_sid_matchup`

```{r id_sid_matchup}
# Import & clean id_sid_matchup2.sas7bdat
id_sid_matchup2 <- read_sas(paste(data_path, "ids", "id_sid_matchup2.sas7bdat", sep = "/"))
id_sid_clean <- id_sid_matchup2 %>% 
  transmute(sid,
            newsid = as.character(newsid),
            shift_num = str_extract(shift_num_ampm, "[:digit:]+"),
            date = d8,
            am_pm = str_extract(shift_num_ampm, "am|pm"),
            shift_start_dttm = startd8time,
            shift_start_tm = ShiftStart,
            shift_end_dttm = ymd_hms(paste(date, ShiftEnd)),
            shift_end_tm = ShiftEnd,
            staff = as_factor(staff))

# preview id_sid_match_cleaned
rmarkdown::paged_table(id_sid_clean)
# Save id_sid_clean to files in csv and Rdata format
# write_csv(id_sid_clean, path = paste(data_path, "Rdata", "id_sid_clean.Rdata", sep = "/"))
# write_csv(id_sid_clean, path = paste(data_path, "csv_files", "id_sid_clean.csv", sep = "/"))
```

## `id_final`

```{r ids_final}
ids_final <- read_sas(paste(data_path, "ids", "finalsids6944.sas7bdat", sep = "/"))
# preview ids_final (no cleaning needed)
rmarkdown::paged_table(ids_final)
# Check if sids repeate within shifts:
ids_final %>% 
  group_by(shift_num) %>% 
  count(sid) %>% 
  filter(n > 1) %>% 
  rmarkdown::paged_table()

```

## `pt_sids`

```{r pt_sids}
pt_sids <- read_sas(paste(data_path, "ids", "patientSIDs_in_completepat.sas7bdat", sep = "/"))
# preview unclean pt_sids
# glimpse(pt_sids)
pt_sid_clean <- pt_sids %>% 
  transmute(sid,
         shift_num = str_extract(shift_num_ampm, "[:digit:]+"),
         shift_ampm = as_factor(str_extract(shift_num_ampm, "am|pm")))
# preview pt_sid_clean
rmarkdown::paged_table(pt_sid_clean)
pt_sid_clean %>% 
  group_by(shift_num) %>% 
  count(shift_ampm) %>% 
  arrange(shift_num) %>% 
  rmarkdown::paged_table()

pt_sid_clean %>% 
  count(shift_ampm) %>% 
  rmarkdown::paged_table()
# Save to file
# write_rds(pt_sid_clean, path = paste(data_path, "Rdata", "pt_sid_clean.Rdata", sep = "/"))
# write_csv(pt_sid_clean, path = paste(data_path, "csv_files", "pt_sid_clean.csv", sep = "/"))
# generate_dictionary(pt_sid_clean)
```

##`staff_sids`

```{r staff_sids}
staff_sids <- read_sas(paste(data_path, "ids", "staffsids_in_eventfile.sas7bdat", sep = "/"))
str(staff_sids)
staff_sids_clean <- staff_sids %>% 
  mutate(across(.cols = !sid, .fns = as.factor)) %>% 
  select(-job)
rmarkdown::paged_table(staff_sids_clean)
staff_sids_clean %>%
  count(ParticipantCat4) %>% 
  rmarkdown::paged_table()
# Save to file
# write_csv(staff_sids_clean, file = paste(data_path, "Rdata", "staff_sids_clean.Rdata", sep = "/"))
# write_csv(staff_sids_clean, file = paste(data_path, "csv_files", "staff_sids_clean.csv", sep = "/"))
# generate_dictionary(staff_sids_clean)
```

## `allstaff`

```{r allstaff_clean}
allstaff <- read_sas(paste(data_path, "staff", "allstaff_numbyshift.sas7bdat", sep = "/"))
# str(allstaff)
allstaff_clean <- allstaff %>% 
  mutate(job_title = as_factor(JOBTITLE),
         .keep = "unused")
# write_rds(allstaff_clean, paste(data_path, "Rdata", "allstaff_clean.Rdata", sep = "/"))
# write_csv(allstaff_clean, paste(data_path, "csv_files", "allstaff_clean.csv", sep = "/"))
rmarkdown::paged_table(allstaff_clean)

# generate_dictionary(allstaff_clean)
```


## `ntab2`  

```{r ntab}
ntab2 <- read_csv(file = paste(data_path, "ids", "nTab2.csv", sep = "/"))
# glimpse(ntab2)
ntab2_clean <- ntab2 %>% 
  transmute(observation = obs,
            sid = sids,
            id = as.character(ID),
            date = dmy(D8),
            shift_num = shiftnum,
            shift_person = person,
            staff = as_factor(staff),
            degree,
            wdegree,
            WtAlpha)

rmarkdown::paged_table(ntab2_clean)
# save to file
# write_rds(ntab2_clean, file = paste(data_path, "Rdata", "ntab2_clean.Rdata", sep = "/"))
# write_csv(ntab2_clean, file = paste(data_path, "csv_files", "ntab2_clean.csv", sep = "/"))
# generate_dictionary(ntab2_clean)
```

ntab2 lists `r max(ntab2_clean$observation)` observations inclusive of all shifts. These observations list up to `r max(ntab2_clean$person)` individuals per shift, all 81 shifts.

## Shift Info

## `shift_times`

```{r shift_times}
shift_times <- read_xls(paste(data_path, "shifts", "Shift Start End times v3.xls", sep = "/"), 
                        col_names = c("date", 
                                      "shift_start_tm", 
                                      "shift_end_tm",
                                      "shift_ampm",
                                      "comments"), 
                        col_types = c("date",
                                      "guess",
                                      "guess",
                                      "text",
                                      "text",
                                      "skip",
                                      "skip",
                                      "skip",
                                      "skip",
                                      "skip",
                                      "skip"), 
                        skip = 1)
# shift_times %>% 
#   str()
rmarkdown::paged_table(shift_times)
# generate_dictionary(shift_times)
# write_csv(shift_times, file = paste(data_path, "csv_files", "shift_times.csv", sep = "/"))
# write_rds(shift_times, file = paste(data_path, "Rdata", "shift_times.Rdata", sep = "/"))
```

## `counts_per_shift`

```{r counts_per_shift}
counts_per_shift <- read_sas(paste(data_path, "shifts", "countspershift.sas7bdat", sep = "/"))
# str(counts_per_shift)
countspershift_clean <- counts_per_shift %>% 
  transmute(shift_num,
         date = d8,
         am_pm = str_extract(shift_num_ampm, "am|pm"),
         staffcount = as.integer(staffcount),
         patcount = as.integer(patcount),
         total = as.integer(total))
rmarkdown::paged_table(countspershift_clean)
# write_rds(countspershift_clean, file = paste(data_path, "Rdata", "countspershift_clean.Rdata", sep = "/"))
# write_csv(countspershift_clean, file = paste(data_path, "csv_files", "countspershift_clean.csv", sep = "/"))
```

## `person_dataset`

```{r person_dataset}
person_dataset <- read_sas(paste(data_path, "patients", "person_dataset.sas7bdat", sep = "/"))

person_dataset <- person_dataset %>% 
  transmute(shift_num = as.integer(shiftnum),
            person = as.character(person),
            staff = as_factor(staff),
            degree,
            minutes)
rmarkdown::paged_table(person_dataset)
# generate_dictionary(person_dataset)
# write_rds(person_dataset, paste(data_path, "Rdata", "person_dataset.Rdata", sep = "/"))
# write_csv(person_dataset, paste(data_path, "csv_files", "person_dataset.csv", sep = "/"))

```

## Room Info

### `location_categories`

```{r location_cats}
location_cats <- read_sas(paste(data_path, "rooms", "location_categories_list.sas7bdat", sep = "/"))
rmarkdown::paged_table(location_cats)
```
### `location_defs`

```{r location_defs}
location_defs <- read_sas(paste(data_path, "rooms", "location_definitions.sas7bdat", sep = "/"))
rmarkdown::paged_table(location_defs)
# generate_dictionary(location_defs)

```

### `room_locs`

```{r room_locs}
room_locs <- read_xlsx(paste(data_path, 
                             "rooms", 
                             "Room Locations and Square Footage - with Correction.xlsx", 
                             sep = "/"), 
                       col_names = c("id", 
                                     "name", 
                                     "type", 
                                     "sqft_area"), 
                       col_types = c("text", 
                                     "text", 
                                     "text", 
                                     "numeric", 
                                     "skip"), 
                       skip = 1)
rmarkdown::paged_table(room_locs)
```
### `locations_list`

```{r locations_list}
locations_list <- read_sas(paste(data_path, 
                                 "rooms", 
                                 "locationslist.sas7bdat", 
                                 sep = "/"))
rmarkdown::paged_table(locations_list)
```


## Patient Data

## Patient Demographics

Demographic information with med_rec numbers and encounter numbers. Note that **med_rec numbers may repeat!**

### `demographics_df` 

```{r demographics_df}
demographic_df <- read_csv(paste(data_path, "patients", "Population_Patient_Demographics.csv", sep = "/"))
# glimpse(demographic_df)
demographics_clean <- demographic_df %>%
  transmute(med_rec_no = as.character(`Medical Record Number`),
         encounter_no = as.character(`Enc .`),
         arrival_date = mdy(`ED Arrival Timestamp`), 
         departure_date = mdy(`ED Departure Timestamp`),
         # across(!where(is.Date)
         #        & !where(is.double)
         #        & !where(is.integer)
         #        & !med_rec_no
         #        & !encounter_no
         #        & !`ED Arrival Timestamp`
         #        & !`ED Departure Timestamp`
         #        & !`Chief Complaint`
         #        & !`% ED Patients Admitted as Inpt or Obs`, 
         #        .fns = as_factor),
         sex = as_factor(Sex, ),
         age = `Patient Age at Visit`,
         race = as_factor(Race),
         chief_complaint = `Chief Complaint`,
         acuity = as_factor(Acuity),
         arrival_mode = as_factor(`Arr Mode`),
         disposition = as_factor(`ED Disposition`)) %>% 
  # str()
# demographics_clean_test %>% 
  mutate(across(where(is.factor), ~recode(.x, `Not Recorded` = NA_character_, `Unknown` = NA_character_))) %>% 
  mutate(acuity = fct_relevel(acuity,
                              "1 Immediate",
                              "2 Emergent",
                              "3 Urgent",
                              "4 Stable",
                              "5 Non Urgent"),
         arrival_cat = fct_recode(arrival_mode,
                                  "EMS" = "EMS Ground",
                                  "EMS" = "EMS Air",
                                  "Custody" = "Police/Jail",
                                  "Ambulatory" = "Private Vehicle",
                                  "Ambulatory" = "Public Trans", 
                                  "Ambulatory" = "Walk"),
         race = fct_relevel(race,
                            "Black",
                            "White",
                            "Hispanic",
                            "Asian",
                            "Other",
                            "1",
                            "2",
                            "3"), .after = 10)
# generate_dictionary(demographics_clean_test)

# generate_dictionary(demographics_clean)

rmarkdown::paged_table(demographics_clean)
demographics_clean %>% 
  count(race) %>% 
  rmarkdown::paged_table()
demographics_clean %>% 
  group_by(acuity) %>% 
  summarise(`Total (n)` = n(), 
            `Black (n)` = sum(race == "Black", na.rm = TRUE),
            `Black (%)` = round((`Black (n)`/`Total (n)`)*100, 2),
            `White (n)` = sum(race == "White", na.rm = TRUE),
            `White (%)` = round((`White (n)`/`Total (n)`)*100, 2),
            `Hispanic (n)` = sum(race == "Hispanic", na.rm = TRUE),
            `Hispanic (%)` = round((`Hispanic (n)`/`Total (n)`) * 100, 2),
            `Other (n)` = sum(race == "Other", na.rm = TRUE),
            `Other (%)` = round((`Other (n)`/`Total (n)`) * 100, 2),
            `Missing (n)` = sum(is.na(race)),
            `Missing (%)` = round((`Missing (n)`/`Total (n)`) * 100, 2)) %>% 
  rename(Acuity = acuity) %>% 
  rmarkdown::paged_table()
# write_rds(demographics_clean, paste(data_path, "Rdata", "demographics_clean.Rdata", sep = "/"))
# write_csv(demographics_clean, paste(data_path, "csv_files", "demographics_clean.csv", sep = "/"))
```

### `pop_pt_demographics`  

```{r pop_pt_demographics}
pop_pt_demographics <- read_csv(paste(data_path, "patients", "Population_Patient_Demographics.csv", sep = "/"))
nms <- c("arrival_date", 
         "med_rec_no", 
         "encounter_no", 
         "sex",
         "age",
         "race",
         "chief_complaint",
         "acuity",
         "arrival_mode", 
         "departure_date",
         "disposition",
         "prop_admitted")
names(pop_pt_demographics) <- nms
# str(pop_pt_demographics)
# generate_dictionary(pop_pt_demographics)
pop_pt_demo_clean <- pop_pt_demographics %>% 
  mutate(med_rec_no = as.character(med_rec_no),
         encounter_no = as.character(encounter_no),
         arrival_date = mdy(arrival_date), .after = 3,
         departure_date = mdy(departure_date),
         sex = as_factor(sex),
         race = as_factor(race),
         acuity = as_factor(acuity),
         arrival_mode = as_factor(arrival_mode)) %>% 
  mutate(across(where(is.factor), ~recode(.x, "Not Recorded" = NA_character_, "Unknown" = NA_character_))) %>% 
  mutate(acuity = fct_relevel(acuity,
                              "1 Immediate",
                              "2 Emergent",
                              "3 Urgent",
                              "4 Stable",
                              "5 Non Urgent"),
         arrival_cat = fct_recode(arrival_mode,
                                  "EMS" = "EMS Ground",
                                  "EMS" = "EMS Air",
                                  "Custody" = "Police/Jail",
                                  "Ambulatory" = "Private Vehicle",
                                  "Ambulatory" = "Public Trans",
                                  "Ambulatory" = "Walk"),
         .after = 9,
         race = fct_relevel(race,
                            "Black",
                            "White",
                            "Hispanic",
                            "Asian",
                            "Other",
                            "1",
                            "2",
                            "3"))

rmarkdown::paged_table(pop_pt_demo_clean)
# generate_dictionary(pop_pt_demo_clean)
# write_csv(pop_pt_demo_clean, file = paste(data_path, "csv_files", "pop_pt_demo_clean.csv", sep = "/"))
# write_rds(pop_pt_demo_clean, file = paste(data_path, "Rdata", "pop_pt_demo_clean.Rdata", sep = "/"))
```

### `pts_population`  

```{r pts_population}
pts_population <- read_sas(paste(data_path, "patients", "Patients_population.sas7bdat", sep = "/"))
rmarkdown::paged_table(pts_population)
# generate_dictionary(pts_population)
# write_csv(pts_population, file = paste(data_path, "csv_files", "pts_population.csv", sep = "/"))
# write_rds(pts_population, file = paste(data_path, "Rdata", "pts_population.Rdata", sep = "/"))
```

### `pts_pop2`

```{r pts_pop2}
pts_pop2 <- read_sas(paste(data_path, "patients", "patients_population2.sas7bdat", sep = "/"))
rmarkdown::paged_table(pts_pop2)
# generate_dictionary(pts_pop2)

# write_csv(pts_pop2, file = paste(data_path, "csv_files", "pts_pop2.csv", sep = "/"))
# write_rds(pts_pop2, file = paste(data_path, "Rdata", "pts_pop2.Rdata", sep = "/"))

```



"completepat.sas7bdat" contains RFID badge locations (by *room number*) for each sampled patient for every second they were monitored in the ED.

The file is large, so subset a random sample of observations with `slice_sample()` to test code for wrangling.

Use `pivot_longer()` to reshape from wide format to long format by collapsing all location-by-second columns (variables `flocX$`, where `X` is the shift time in seconds that the respective SID was identified in room number `flocX`) into two columns, names set to `seconds` and values to `location`. 

> Note that `pivot_longer()` will create *many duplicate SIDs and many* `NA`s which will require filtering.

## Patient RFID Data  

### `pt_complete`

```{r pt_complete}
#| code-fold: true
pt_complete <- read_sas(paste(data_path, "patients", "completepat.sas7bdat", sep = "/"))
# glimpse(pt_complete)
# 1. subset a random sample of 30 observations for data transformation code preparation
# set.seed(711)
# pt_complete_sample <- pt_complete %>% slice_sample(n = 30)
# pt_complete_sample %>%
pt_complete_long <- pt_complete %>%

# 2. Pivot the data.frame from wide to long by placing all column names that start with "floc" into a new column, "seconds," and placing respective observations for each "floc" variable into a "location_num" column
  pivot_longer(cols = starts_with("floc"), 
               names_to = "shift_second", # moves column names to new column "seconds"
               values_to = "location_num", # moves location numbers to new column "location_num"
               values_drop_na = TRUE) %>% # removes all rows (observations) without location data
# 3. Delete the prefix "floc" from `seconds` and change class to numeric
  mutate(shift_second = as.double(str_replace(shift_second, "floc", "")),
         location_num = as_factor(location_num),
         shift_num_ampm = str_trim(shift_num_ampm),
         am_pm = as_factor(str_extract(shift_num_ampm, "am|pm")),
         shift_num = as.integer(str_extract(shift_num_ampm, "[:digit:]+")),
         date = make_date(year = year, month = mon, day = day)) %>% 
  select(sid, date, shift_num, am_pm, location_num, shift_second) %>% 
  # count(am_pm)
  mutate(day_night = as_factor(if_else(am_pm == "am", "day", "night")),
         .after = 4) # %>% 
  # glimpse()
# Save as .Rdata file
rmarkdown::paged_table(pt_complete_long)
# write_rds(pt_complete_long, file = paste(data_path, "Rdata", "pt_complete_long.Rdata", sep = "/"))
# write_csv(pt_complete_long, file = paste(data_path, "csv_files", "pt_complete_long.csv", sep = "/"))
# generate_dictionary(pt_complete_long)
```

> After transforming & cleaning "completepat.sas7bdat", `pt_complete_long` is __`r ncol(pt_complete_long)`__ variables wide and __`r nrow(pt_complete_long)`__ rows long.  


## Staff Data

## Staff RFID data


The SAS data file, "completestaff.sas7bdat," contains RFID badge location (by *room number*) for all staff each second of every shift.

### `staff_complete`

```{r staff_complete}
# 1. pull "completestaff.sas7bdat" from `sas_data_list` as `tibble()` 
staff_complete <- read_sas(paste(data_path, "staff", "completestaff.sas7bdat", sep = "/"))
# Subset a random sample of 30 observations
# set.seed(321)
# staff_complete_sample <- staff_complete %>% slice_sample(n = 30)
# 2. Use a random sample of staff_complete for data cleaning & transformation 
staff_complete_long <- staff_complete %>% 
  # 2a. Pivot the data.frame from wide to long by placing all column names that start with "floc" into a new column, "seconds," and placing respective observations for each "floc" variable into a "location_num" column
  pivot_longer(cols = starts_with("floc"), 
               names_to = "seconds", 
               values_to = "location_num",
               values_drop_na = TRUE) %>% 
  # 2b. Remove the prefix "floc" from `time_seconds` and keep the digits as `seconds`
  mutate(seconds = as.double(str_replace(seconds, "floc", "")),
         location_num = as_factor(location_num),
         shift_num_ampm = str_trim(shift_num_ampm),
         shift_num = as.integer(str_extract(shift_num_ampm, "[:digit:]+")),
         am_pm = as_factor(str_extract(shift_num_ampm, "am|pm")),
         date = make_date(year = year, month = mon, day = day)) %>%
  select(-d8:-mon)

# staff_complete_sample2
# write_rds(staff_complete_long, file = paste(data_path, "Rdata", "staff_complete_long.Rdata", sep = "/"))
# write_csv(staff_complete_long, file = paste(data_path, "csv_files", "staff_complete_long.csv", sep = "/"))
# 3. View data frame structure
# generate_dictionary(staff_complete_long)
rmarkdown::paged_table(staff_complete_long)
```




## Network Data  

## Whole Networks

### `net_all_xl`


```{r net_all_xl}
#| label: "net_all"
net_all_xl <- read_xlsx(paste(data_path, "networks", "network_allshifts.xlsx", sep = "/"))
str(net_all_xl)
net_all_xl %>% 
  rmarkdown::paged_table()
```
### `net_all_sas`

```{r net_all_sas}
net_all_sas <- read_sas(paste(data_path, "networks", "network_allshifts.sas7bdat", sep = "/"))
# str(net_all_sas)
net_all_sas %>% 
  rmarkdown::paged_table()
generate_dictionary(net_all_sas)
```


## Edges 

### `edges_all`

```{r edges_all}
#| label: "edges_all"
edges_all <- read_sas(paste(data_path, "networks", "allshifts_edges.sas7bdat", sep = "/"))
# str(edges_all)
edges_all %>% 
  rmarkdown::paged_table()
```

```{r edges2}
#| label: "edges2"
edges2 <- read_sas(paste(data_path, "networks", "edges2.sas7bdat", sep = "/"))
rmarkdown::paged_table(edges2)
```
## Events

### `events_long`

```{r events_long}
events_long <- read_sas(paste(data_path, "events", "eventfile_long2.sas7bdat", sep = "/"))
rmarkdown::paged_table(events_long)
# write_csv(events_long, file = paste(data_path, "csv_files", "events_long.csv", sep = "/"))
```

`events_long` has duplicates, use `eventfile_nodup` if needed.

### `events_unique`

```{r events_unique}
events_unique <- read_sas(paste(data_path, "events", "events_uniquepersid2.sas7bdat", sep = "/"))
rmarkdown::paged_table(events_unique)
```

### `eventfile_nodup`

```{r eventfile_nodup}
eventfile_nodup <- read_sas(paste(data_path, "events", "eventfile_long6944_noduploc.sas7bdat", sep = "/"))
eventfile_nodup %>% 
  rmarkdown::paged_table()
# write_csv(eventfile_nodup, file = paste(data_path, "csv_files", "eventfile_nodup.csv", sep = "/"))
# write_rds(eventfile_nodup, file = paste(data_path, "Rdata", "eventfile_nodup.Rdata", sep = "/"))
```


## Fully Joined Datasets  

```{r full_join}
rdata_dir <- paste0(getwd(), "/Data/Rdata/")

# population.Rdata
population <- read_rds(paste0(rdata_dir, "population.Rdata"))
glimpse(population)
# names(population)
# demographics.Rdata
# demographics <- read_rds(paste0(rdata_dir, "demographics_clean.Rdata"))
# glimpse(demographics)

# pts_all.Rdata
pts_all <- read_rds(paste0(rdata_dir, "pts_all.Rdata"))
# glimpse(pts_all)
# names(pts_all)
test <- population %>%
  full_join(pts_all, by = c("encounter_no",
                            "sex",
                            "age",
                            "race",
                            "chief_complaint",
                            "acuity",
                            "arrival_mode",
                            "arrival_cat",
                            "disposition"))
glimpse(test)
all_pts <-  test %>%
  unnest(cols = "nested_loc", keep_empty = TRUE) # %>%
  # glimpse()
# pts_all == pts_nested
glimpse(all_pts)



# staff_all == staff_locs_nested
staff_all <- read_rds(paste0(rdata_dir, "staff_all.Rdata"))
# glimpse(staff_all)

all_staff <- staff_all %>%
  unnest(cols = nested_loc, keep_empty = TRUE)
glimpse(all_staff)

# what file format do data need to for uploading to google cloud???
# write_rds(all_pts, file = "all_pts.Rdata")
# write_rds(all_staff, file = "all_staff.Rdata")

```


```{r wrangling_old_code, eval=FALSE}



# directory path
rdata_dir <- "/Users/tommy-two/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data/Rdata"
rdata_archive <- paste0(rdata_dir, "/rds_archive")
# rdata_dir
# WRANGLING----

## X--ntab2_final----
ntab2_final <- read_rds(paste0(rdata_dir, "/ntab2_final.Rdata"))
# view loaded data
ntab2_final %>%
  glimpse()
# ntab2_clean <- ntab2_clean %>%
#   mutate(shift_num = as.character(shift_num),
#          shift_person = as.character(shift_person),
#          sid = format(sid, scientific = FALSE))
# ntab2_clean <- ntab2_clean %>%
#   relocate(shift_person, .after = id) %>%
#   mutate(sid = str_trim(sid))
# ntab2_nms <- names(ntab2_clean)

# Duplicate sid for two patients in the same shift (shift 91; 10/07/2009):
# ntab2_final %>%
#   mutate(sid = (sid)) %>%
#   group_by(shift_num) %>%
#   count(sid) %>%
#   filter(n > 1)
# one duplicate sid from shift 91:
# ntab2_final %>%
#   filter(shift_num == "91" & sid == "2.17E+04")
# id_sid_final %>%
#   glimpse()
# ntab2_final %>%
#   # filter(str_detect(sid, "^2")) %>%
#   filter(shift_num == "8") %>%
#   count(sid) %>%
#   arrange(desc(n)) #%>%
#   filter(str_detect(sid, "2\\.17"))
# id_sid_final %>%
#   # glimpse()
#   group_by(shift_num) %>%
#   count(sid) %>%
#   arrange(desc(n)) %>%
#   filter(str_detect(sid, "217"))
#   filter(str_detect(sid, "217"))
#   mutate(staff = if_else(staff == 1, "Staff", "Patient"))
# write_rds(ntab2_clean, file = paste0(rdata_dir, "/ntab2_final.Rdata"))
# rm(ntab2_clean)

## X--id_sid_clean----
id_sid_clean <- read_rds(paste0(rdata_dir, "/id_sid_clean.Rdata"))
# view loaded data
# id_sid_clean %>%
#   glimpse()
# compare to original file: "id_sid_matchup2.sas7bdat"
# idsid_matchup <- read_sas("Data/ids/id_sid_matchup2.sas7bdat")
id_sid_clean %>%
  glimpse()
# idsid_matchup %>%
  # glimpse()
  # group_by(numshift) %>%
  # count(newsid) %>%
  # filter(n > 1)
  # summarise(max(newsid))

id_sid_clean <- id_sid_clean %>%
  rename(shift_ampm = am_pm,
         shift_person = newsid) # %>%
# glimpse()
# write_rds(id_sid_clean, paste0(rdata_dir, "/id_sid_final.Rdata"))
# rm(id_sid_clean)

id_sid_clean %>%
  group_by(shift_num) %>%
  count(sid) %>%
  arrange(desc(n))
# **ids_final ------
id_sid_final <- read_rds(paste0(rdata_dir, "/id_sid_final.Rdata"))
id_sid_final %>%
  group_by(staff) %>%
  count(shift_num) %>%
  ungroup() %>%
  pivot_wider(names_from = staff, values_from = n) %>%
  transmute(`Shift Number` = as.numeric(shift_num),
            `Patients (n)` = `0`,
            `Staff (n)` = `1`) %>%
  arrange(`Shift Number`) %>%
  rownames_to_column() %>%
  mutate(`Shift Number` = rowname,
         .keep = "unused") %>%
  relocate(`Shift Number`, `Patients (n)`, `Staff (n)`) %>%
  mutate(`Total (n)` = `Patients (n)` + `Staff (n)`)


id_sid_final %>%
  glimpse()
## **staff_sids----
# staff_sids <- read_csv(file = "Data/csv_files/staff_sids_clean.csv",
#                        col_names = c("sid", "job_sah", "participant_cat", "jobtitle"),
#                        col_types = "cfff",
#                        skip = 1)
# Read saved R object from file = "staff_sids_clean.Rdata"
staff_sids <- read_rds(paste0(rdata_dir, "/staff_sids-5_1_22.Rdata"))
staff_sids %>%
  # glimpse()
  count(sid) %>%
  arrange(desc(n))
# write_rds(staff_sids, paste0(rdata_dir, "/staff_sids_final.Rdata"))
# rm(staff_sids)

# no duplicates in staff sid. 88 total staff members.
staff_sids_final <- read_rds(paste0(rdata_dir, "/staff_sids_final.Rdata"))
staff_sids_final %>%
  glimpse()

## **pt_sids----
pt_sid_clean <- read_rds(paste0(rdata_dir, "/pt_sid_clean.Rdata"))
# view loaded data
shifts_vec <- pt_sid_clean %>%
  mutate(shift_num = as.integer(shift_num)) %>%
  arrange(shift_num) %>%
  pull(shift_num) %>%
  unique() %>%
  as.character()
# set seed for reproducibility:
set.seed(711)
# select a random sample of
random_shifts <- sample(shifts_vec, 20)

# many duplicate sids:
pt_sid_clean %>%
  count(sid) %>%
  filter(n > 1) %>%
  arrange(desc(n))
# no duplicate sids when grouped by shift:
pt_sid_clean %>%
  group_by(shift_num) %>%
  count(sid) %>%
  arrange(desc(n))
# write_rds(pt_sid_clean, paste0(rdata_dir, "/pt_sid_final.Rdata"))
# rm(pt_sid_clean)


## **population (from demographics_clean)----
demographics_clean <- read_rds(paste0(rdata_dir, "/demographics_clean.Rdata"))
demographics_clean %>%
  glimpse()
# saved as population.Rdata
# write_rds(demographics_clean, paste0(rdata_dir, "/population.Rdata"))

## **pts_pop_clean----
pts_population <- read_rds(paste0(rdata_archive, "/pts_population.Rdata"))
pts_pop_clean <- pts_population %>%
  mutate(sid,
         encounter_no = str_trim(EncounterNum),
         encounter_no2 = as.character(encnum),
         pt_id = as.character(ID),
         date = d8,
         shift_num = as.character(numshift),
         shift_ampm = factor(shift_ampm),
         sex = factor(Sex),
         age = as.numeric(Patient_Age_at_Visit),
         age2 = AGE,
         race = factor(Race,
                       levels = c("Black",
                                  "White",
                                  "Hispanic",
                                  "Other"),
                       exclude = c("Not Recorded", "1", "2", "3")),
         chief_complaint = Chief_Complaint,
         acuity = factor(Acuity,
                         levels = c("1 Immediate",
                                    "2 Emergent",
                                    "3 Urgent",
                                    "4 Stable",
                                    "5 Non Urgent"),
                         exclude = "Not Recorded"),
         arrival_dttm = ED_Arrival_Timestamp,
         arrival_tm = timeARR,
         arrival_mode = factor(Arr_Mode, exclude = "Not Recorded"),
         arrival_cat = fct_recode(arrival_mode,
                                  "EMS" = "EMS Ground",
                                  "EMS" = "EMS Air",
                                  "Custody" = "Police/Jail",
                                  "Ambulatory" = "Private Vehicle",
                                  "Ambulatory" = "Public Trans",
                                  "Ambulatory" = "Walk"),
         departure_dttm = ED_Departure_Timestamp,
         departure_tm = timeDEP,
         time2tag = duration_UntilTag,
         start_date = startd8time,
         end_date = endd8time,
         los = durationInED,
         los_minutes = MinutesInED,
         los_hours = hrsinED,
         disposition = ED_Disposition,
         job = JOB,
         approached = factor(approached),
         participant = factor(participantyn),
         pt_with_data = factor(Pt_withData),
         # levels = c("Yes" = 1,
         #            "No" = 0)),
         participant_final = factor(Participant_final),
         .keep = "none") %>%
  mutate(across(where(is.factor), ~na_if(., ""))) %>%
  mutate(across(where(is.character), ~na_if(., ""))) %>%
  mutate(across(where(is.factor), ~na_if(., "Not Recorded"))) %>%
  mutate(across(where(is.character), ~na_if(., "Not Recorded"))) %>%
  relocate(shift_ampm, .after = date) %>%
  relocate(arrival_cat, .after = arrival_mode) %>%
  relocate(approached, .before = participant)

# check results
pts_pop_clean %>%
  glimpse()
  # filter(encounter_no %in% demographics_clean$encounter_no)
  count(shift_num)

  # filter(encounter_no %in% pts_pop_clean$encounter_no)
# write pts_pop to rds file:
# write_rds(pts_pop, paste0(rdata_dir, "/pts_pop_clean.Rdata"))
# rm(pts_pop)

## X--allstaff_clean----
allstaff_clean <- read_rds(paste0(rdata_dir, "/allstaff_clean.Rdata"))
allstaff_clean %>%
  glimpse()

## **pt_complete----
# pt_complete <- read_rds(paste0(rdata_dir, "/pt_complete_long.Rdata"))
# view loaded data
# pt_comp <- pt_complete %>%
#   mutate(shift_num = as.character(shift_num),
#          day_night = NULL) %>%
#   rename(shift_ampm = am_pm)
pt_comp_nested <- read_rds(paste0(rdata_dir, "/pt_comp_nested-5_2_22.Rdata"))
# create unnested version
pt_comp_unnested <- pt_comp_nested %>%
  unnest(cols = c(data))
# view unnested version
glimpse(pt_comp_unnested)
# ungroup unnested df
pt_comp_unnested %>%
  ungroup() %>%
  count(date) %>%
  arrange(date) %>%
  plot(date, n)
# renest df by sid & shift_num only
pt_comp_nested <- pt_comp_unnested %>%
  group_by(sid, shift_num) %>%
  nest(nested_loc = c(location_num, shift_second)) # %>%
# check for duplicates:
# pt_comp has many duplicate sids:
pt_comp %>%
  count(sid) %>%
  filter(n > 1)

# pt_comp_nested has no duplicate sids
  pt_comp_nested %>%
    # glimpse() %>%
    count(sid) %>%
    filter(n > 1)
# no duplicates found
# glimpse(pt_comp_unnested)
glimpse(pt_comp_nested)
# write_rds(pt_comp_nested, file = paste0(rdata_dir, "/pt_locs_nested.Rdata"))
# rm(pt_comp_nested)

## **staff_complete----
staff_complete <- read_rds(paste0(rdata_dir, "/staff_complete_long.Rdata"))

staff_comp <- staff_complete %>%
  mutate(shift_num = as.character(shift_num),
         firstday = NULL) %>%
  rename(shift_ampm = am_pm,
         shift_second = seconds) %>%
  relocate(sid, date, shift_num, shift_ampm, location_num, shift_second)
# exclude from staff_sids all rows with sid not in staff_complete
stf_comp_sids <- staff_complete %>%
  distinct(sid)
stf_sids_sids <- staff_sids %>%
  select(sid)
excluded_staff_sids <- staff_sids %>%
  filter(!(stf_sids_sids$sid %in% stf_comp_sids$sid)) %>%
  pull(sid)
staff_sids_88 <- staff_sids %>%
  filter(!(sid %in% excluded_staff_sids))

# write_rds(staff_sids_88, file = paste0(rdata_dir, "/staff_sids-5_1_22.Rdata"))
# write_csv(staff_sids_88, file = paste0("Data/csv_files/staff_sids-5_1_22"))
staff_comp %>%
  glimpse()

# staff_comp has many duplicated sid numbers, but not within shifts:

staff_comp %>%
  count(sid) %>%
  filter(n > 1)



staff_comp_nested <- staff_comp %>%
  group_by(sid, shift_num) %>%
  nest(nested_loc = c(location_num, shift_second)) # %>%
# staff_comp_nested has no duplicated sids
# the following table shows number of observed staff (n) per shift, totals of staff working per shift is not available
staff_comp_nested %>%
  ungroup() %>%
  count(shift_num) %>%
  arrange(desc(n))
  # filter(n > 1)
# to unnest staff_comp_nested, use:
staff_comp_nested %>%
  unnest(cols = c(nested_loc))
# view glimpse() of staff_comp_nested:
staff_comp_nested %>%
  glimpse()
write_rds(staff_comp_nested, paste0(rdata_dir, "/staff_locs_nested.Rdata"))
rm(staff_comp, staff_comp_nested)
rm(staff_complete)
## **events_clean----
# load eventfile_nodup.Rdata
events <- read_rds(paste0(rdata_dir, "/eventfile_nodup.Rdata"))

# number of staff: 88
# events %>%
#   filter(ParticipantCat2 == "STAFF") %>%
#   group_by(sid) %>%
#   count() %>%
#   ungroup() %>%
#   summarise(n_staff = length(sid))



# clean events:
events_clean <- events %>%
  mutate(sid = sid,
         pt_id = patient,
         encounter_no = str_trim(EncounterNum),
         participant_cat = factor(ParticipantCat4),
         job_cat = factor(job),
         job_name = if_else(jobfullname == "", "Patient", jobfullname),
         date = ymd(d8),
         shift_num = shift_num,
         shift_start_dttm = startd8time,
         shift_end_dttm = endd8time,
         tag_on_dttm = tag_starttime,
         tag_off_dttm = tagofftime,
         event_id = factor(event),
         event_dttm = ymd_hms(et),
         event_tm = ti,
         room_index = locindex,
         room_name = room,
         .keep = "none",
         .after = "sid") %>%
  relocate(pt_id, date, shift_num, participant_cat, .after = encounter_no)

# write_rds(events_clean, paste0(rdata_dir, "/events_final.Rdata"))
# rm(events_clean, events)
pt_id_vec <- events_clean %>%
  filter(pt_id %in% pts_pop_clean$id_no) %>%
  pull(pt_id)
# c(pt_id_vec)
pt_id_vec <- pt_id_vec %>%
  # group_by(pt_id) %>%
  distinct(pt_id) %>%
  pull(pt_id)

pts_pop_clean %>%
  filter(id_no %in% pt_id_vec$pt_id)

events_clean %>%
  filter(pt_id %in% pt_id_vec$pt_id)

events_cl_nested <- events_clean %>%
  group_by(shift_num, sid, encounter_no, date) %>%
  nest()


# JOINING DFs:-----
# Read saved data objects:
id_sid <- read_rds(paste0(rdata_dir, "/id_sid_final.Rdata"))
ntab2 <- read_rds(paste0(rdata_dir, "/ntab2_final.Rdata"))
pt_info <- read_rds(paste0(rdata_dir, "/pt_info_final.Rdata"))
pt_sids <- read_rds(paste0(rdata_dir, "/pt_sid_final.Rdata"))
pts_nested <- read_rds(paste0(rdata_dir, "/pt_locs_nested.Rdata"))
staff_nested <- read_rds(paste0(rdata_dir, "/staff_locs_nested.Rdata"))
staff_sids <- read_rds(paste0(rdata_dir, "/staff_sids_final.Rdata"))
events <- read_rds(paste0(rdata_dir, "/events_final.Rdata"))

## staff_all <-  staff_nested + staff_sids =====
staff_all <- staff_nested %>%
  left_join(staff_sids, by = "sid")

staff_all %>%
  glimpse()



# write_rds(staff_all, paste0(rdata_dir, "/staff_all.Rdata"))

# pts_all <- pts_nested + pt_info----

pt_info %>%
  glimpse()
# names(pts_nested)
# names(pt_info)
# test <- pts_nested %>%
#   full_join(pt_info, by = c("sid",
#                             "date",
#                             "shift_num",
#                             "shift_ampm"))

# glimpse(test)
# pts_nested %>%
  # glimpse()
# move list column to last position
pts_all <- pts_all %>%
  ungroup() %>%
  relocate(nested_loc, .after = participant_final)
pts_all %>%
  glimpse()

# pts_all_unnested <- pts_all %>%
#   unnest(cols = nested_loc) #%>%

# save pts_all_unnested as .csv & .Rdata files:
# write_csv(pts_all_unnested, paste0("Data/csv_files/pts_all_unnested.csv"))
# write_rds(pts_all, paste0(rdata_dir, "/pts_all.Rdata"))






# SLICES:----

sliced_pts <- pts_all %>%
  unnest(cols = c(nested_loc)) %>%
  group_by(sid, shift_num) %>%
  slice_sample(prop = 0.005)

# sliced_pts %>%
#   ungroup() %>%
#   group_by(encounter_no)# %>%
# summarise(n_pts = n(),
#           n_participants = sum(as.numeric(participant_final), na.rm = TRUE))


# sliced_pts %>%
#   # glimpse()
#   ungroup() %>%
#   count(sid) %>%
#   summarise(sliced_pts_n = n(), sliced_obs_n = sum(n))
# summarise(sum(n))
```




# EXPLORATION  

```{r exploratory_old_code, eval=FALSE}
# SETUP ----
## create path to data archive
rdata_dir <- "/Users/tommy-two/Documents/1_Research/2_Data_Science/0_Projects/1_NACI/Data/Rdata"
rdata_archive <- paste0(rdata_dir, "/rds_archive")
## load packages
# library(naniar) # for missing data
# library(lubridate)
library(labelled)
library(vtable)# good package for summary tables
library(tidyverse)
## import data sets
population_730 <- read_rds(paste0(rdata_dir, "/population.Rdata"))
n_patients_730 <- population_730 %>%
  nrow()
# n_patients_730
population_81 <- read_rds(paste0(rdata_dir, "/pt_info_final.Rdata"))
n_patients_81 <- population_81 %>%
  nrow()
# n_patients_81

# Make a table of patient & staff counts (n) per shift (shift_counts_table.csv)----
id_sid_final <- read_rds(paste0(rdata_dir, "/id_sid_final.Rdata"))
shift_counts_table <- id_sid_final %>%
  group_by(staff) %>%
  count(shift_num) %>%
  ungroup() %>%
  pivot_wider(names_from = staff, values_from = n) %>%
  transmute(`Shift Number` = as.numeric(shift_num),
            `Patients (n)` = `0`,
            `Staff (n)` = `1`) %>%
  arrange(`Shift Number`) %>%
  rownames_to_column() %>%
  mutate(`Shift Number` = rowname,
         .keep = "unused") %>%
  relocate(`Shift Number`, `Patients (n)`, `Staff (n)`) %>%
  mutate(`Total (n)` = `Patients (n)` + `Staff (n)`)
# write_csv(shift_counts_table, paste0(getwd(), "/Tables/shift_counts_table.csv"))




```

# SessionInfo

```{r SessionInfo}
sessioninfo::session_info()%>%
  details::details(
    summary = 'Current session info',
    open    = FALSE
  )
```
